---
title: "Figures output"
author: ""
date: "`r Sys.Date()`"
output:
  pdf_document:
    keep_tex: false
    latex_engine: xelatex
    fig_caption: yes
    highlight: haddock
      
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo =F,message = F,warning = F,include = F)
```


# RNAseq data loading
```{r,include=F}
rm(list=ls())
Path=getwd()
COUNT_path=file.path(Path,"data" ,"RNAseq","Count")
Unused_samples=c("JH-2-084","JH-2-106") # These samples are not inclucded in the analyssi
Reference_path="/Users/lyu.yang/Tools/R/Public_database/Reference/Human_GSEA" # Download GMT files from  from https://www.gsea-msigdb.org/gsea/msigdb
Process_data=file.path(Path,"data","RNAseq","Process_data")

dir.create(COUNT_path,recursive = T)
dir(Reference_path)
```


```{r,include=F}
library(dplyr)
library(ggplot2)
library("ggbreak")
library(synapser)
library(synapserutils)
```


```{r}
# color_setting

if (!file.exists(file.path(COUNT_path,"Biobank-samples-map-clinical-102523.xlsx"))){
  synLogin()


# DOWNLOAD CLINICAL METADATA (uncomment next line to download file)
 synapser::synGet("syn52798607", downloadLocation = COUNT_path)
}



# READ IN OFFLINE SAMPLESHEET
jhu_clinicaldata_1 <- readxl::read_xlsx(file.path(COUNT_path,"Biobank-samples-map-clinical-102523.xlsx")) %>% 
  dplyr::filter(assay == "Whole Exome Sequencing") %>%  #select WES samples
  dplyr::filter(batch != "WU_batch3") %>%  # select only batch1, 2, 3 samples
  dplyr::filter(tissue == "primary tumor") %>%  # select only tumor samples
  dplyr::mutate(Aliquot_ID_Synapse_NEW = tidyr::replace_na(Aliquot_ID_Synapse_NEW, "A")) %>% 
  dplyr::mutate(Tumor_Sample_Barcode = stringr::str_c(SpecimenID_Synapse_NEW, "-", Aliquot_ID_Synapse_NEW)) %>% 
  dplyr::mutate(
    # Create categories
    age_group = dplyr::case_when(
      age <= 10            ~ "0-10",
      age > 10 & age <= 20 ~ "11-20",
      age > 20 & age <= 30 ~ "21-30",
      age > 30 & age <= 40 ~ "31-40",
      age > 40 & age <= 50 ~ "41-50",
      age > 50             ~ "> 50"
    ),
    # Convert to factor
    age_group = factor(
      age_group,
      level = c("0-10", "11-20","21-30", "31-40", "41-50", "> 50")
    )
  ) %>% 
  dplyr::select(tumorType, sex, age_group, location, tissue, individualID, Tumor_Sample_Barcode, NF1_loss_clinical_sequencing, NF1_loss_clinical_diagnosis) %>% 
  dplyr::filter(!(Tumor_Sample_Barcode %in% c("JH-2-015-6AD96-A", "JH-2-106-B73E8-A", "JH-2-106-1447G-A", "JH-2-084-F9A84-A", "JH-2-084-AE169-A"))) %>% #remove 2-015TD, JH-2-106, JH-2-084 
  purrr::set_names(nm = c(
    "tumorType", 
    "sex", 
    "age_group", 
    "location", 
    "tissue", 
    "individualID", 
    "Tumor_Sample_Barcode", 
    "NF1_clinical_sequencing", 
    "NF1_clinical_diagnosis"
  ))

```

## Assign colors to the metadata for plotting 

```{r colors, echo=TRUE, eval=TRUE, results='hide', message=FALSE, warning=FALSE}

# tumortype_colors <- c("#0D3B66", "#2d728f","#ab3428","#61C9A8","#7E52A0", "#F4D35E","#f49e4c")
tumortype_colors <- c("#0D3B66", "#2d728f","#ab3428","#61C9A8","#7E52A0", "#F4D35E")
names(tumortype_colors) <- jhu_clinicaldata_1$tumorType %>%unique()

```


```{r}
load(file.path(Process_data,"MPNST_VS_PN_Deseq2_outut.rdata"))

```

```{r}
#load(file.path(Reference_path,"Human_GSEA","GO_KEGGP_WP_C2CGP_Reactome_pathway_list.rdata"))

PCA_Plot_3<- function (data,Annotation,VAR,Color,manu_color) {
  # logcountdata row:genes,column: samples
  pca <- prcomp(data) 
  pca_out<-as.data.frame(pca$x)
  df_out<- pca_out %>%tibble::rownames_to_column(var=VAR)  %>% left_join(., Annotation) 
  #df_out<- merge (pca_out,Annotation,by.x=0,by.y=0) 
  PC1_var=summary(pca)$importance%>%as.data.frame()%>%.[2,1]%>%scales:: percent(accuracy = 0.01)
  PC2_var=summary(pca)$importance%>%as.data.frame()%>%.[2,2]%>%scales:: percent(accuracy = 0.01)
  # label_color<- factor(df_out[,group])
  ggplot(df_out,aes_string(x="PC1",y="PC2")) +geom_point(aes_string(colour = Color),size=3) +scale_color_manual(values = manu_color)+xlab(paste0("PC1(",PC1_var,")"))+ylab(paste0("PC2(",PC2_var,")"))+theme(legend.position = "right",legend.direction = "vertical",legend.title = element_text(size=12, vjust = .5, hjust = .3))
}

# Batch_count_plot=function(logcount,sample_anno){
#   
#   logcount_2=logcount%>%t() %>%as.data.frame() %>%mutate(Sample_name=rownames(.))
#   Level=sample_anno %>%arrange(batch)%>%.$Sample_name
#   data_mod = reshape2::melt(logcount_2, id.vars="Sample_name", 
#                             measure.vars=rownames(logcount)) %>%left_join(sample_anno)%>%arrange(batch)
#   
#   data_mod$Sample_name=factor(data_mod$Sample_name,levels =Level )
#   #return(logcount_2)
#  ggplot(data_mod)+geom_boxplot(aes(x=Sample_name, y=value, color=batch))+xlab("Sample")+ylab("log2(Count+1)")+theme(axis.text.x=element_blank())
#   
#   
# }


pheatmap_pathway_3<- function (expression_data, Gene_select,pathway_name,Annotation,Label_def,anno_color) {
 # Gene_select<-Pathway_list[,pathway_name]%>%unique()
  Gene_select_anno= expression_data[,colnames(expression_data) %in% Gene_select] %>%t()
 
  # Anno_expression_data=Gene_select_anno[,c("SYMBOL",Group_select)] %>% as.data.frame() %>% distinct() %>% na.omit()
  # rownames(Anno_expression_data)=Anno_expression_data[,"SYMBOL"]
  # Annotation=group_anno["Gene_type"]
  # input= Anno_expression_data[,Group_select]
  # F2_pheatmap <- pheatmap::pheatmap(input, cellwigermline calling GATKdth = 10, cellheight = 12, scale = "row",
  #                                   treeheight_row = 5,
  #                                   show_rownames = T,show_colnames = T, 
  #                                   annotation_col= Annotation,
  #                                   # annotation_row=Annotation,
  #                                   annotation_legend=Label_def,
  #                                   cluster_rows = T,  cluster_cols = F,clustering_distance_rows = "euclidean")
  pheatmap::pheatmap(Gene_select_anno, cellwigermline= 2, cellheight =10, scale = "row",
                     treeheight_row = 5,main=pathway_name,
                     show_rownames = T,show_colnames = F,
                     annotation_col= Annotation,
                     annotation_colors= anno_color,
                     # annotation_row=Annotation,
                     #annotation_legend=Label_def,
                     cluster_rows = T,  cluster_cols = F,clustering_distance_rows = "euclidean",silent = F)
  
}

singcore_Normalized_GSEA=function (data,genelist, is.na=T) {
  nm=names(genelist)
  plist=list()
  Data_prepare=singscore::rankGenes (data)
 # n_genes=length(intersect(genelist,rownames(data)))
  
   if( is.list(genelist)) {
    for (i in 1:length(genelist)) { 
      n_genes=length(intersect(unique(genelist[[i]]),rownames(data)))
      if (n_genes>4){
      scoredf<- singscore::simpleScore (Data_prepare, upSet=unique(genelist[[i]]))%>%.[1]
      scoredf$Normalized=scale(scoredf[,1])
      colnames(scoredf)=c(paste0(nm[i],"_singscore") , paste0(nm[i],"_Normalized_singscore" ))
      if(nrow(scoredf)==ncol(data)) {
        plist[[nm[i]]]=scoredf
      }
      }
     }
  }
  if(is.data.frame(genelist))  {
    
  for (i in seq_along(genelist)) { 
    n_genes=length(intersect(unique(genelist[[i]]),rownames(data)))
    if (n_genes>4){
    scoredf<- singscore::simpleScore (Data_prepare, upSet=unique(genelist[,i]))%>%.[1]
    scoredf$Normalized=scale(scoredf[,1])
    colnames(scoredf)=c(paste0(nm[i],"_singscore") , paste0(nm[i],"_Normalized_singscore" ))
    if(nrow(scoredf)==ncol(data)) {
      plist[[nm[i]]]=scoredf
     }
    }
  }
  }
  return(plist)
  #output=do.call("cbind",plist)
}

Singscore_data_plot_2=function(gene_list,expression_data,expression_data_anno,Annotation_File,stat_method,Anno_color){
  Gene_singscore= singcore_Normalized_GSEA(expression_data,gene_list)
  Manu_color=unlist(Anno_color )
  names(Manu_color)=gsub("Group.", "",names(Manu_color))

  Annotation_file=Annotation_File[,c("Group")]  %>%as.data.frame()
  rownames(Annotation_file)=Annotation_File$Sample_name
  plist=list()
  plist[[2]]=GGplot_singscore_boxplot_2(Gene_singscore[[1]],Annotation_File,"Group","Sample_name",Method=stat_method,Manu_color)
  
  plist[[1]]=pheatmap_pathway_3 (expression_data_anno,unlist(gene_list),names(gene_list), Annotation_file,"V1",Anno_color)
  return(plist)
}

GGplot_singscore_boxplot_2=function(singscore,Annotation,group_col,Var,Method,manu_color) {
  library(ggplot2)
  library(ggpubr)
  gene_name=colnames(singscore)[2]
  Data=singscore%>%tibble::rownames_to_column(.,var=Var)
  Data2= Data %>% left_join(.,Annotation,by=Var)%>%as.data.frame()


  ggplot(Data2,aes_string(x="Group",y=gene_name,fill=group_col)) +geom_boxplot() +xlab("")+
    ylab(gene_name)+scale_fill_manual(values = manu_color)+ theme(axis.title.y = element_text(size=8),axis.text.x = element_text(angle=90, vjust=.5, hjust=1)) +stat_compare_means(method = Method)



}

Valcano_plot_rnaseq=function(Data,Pathway_list,pathway_name){
  pathway_genes=Pathway_list [[pathway_name]]%>%as.character()
  
  library(ggrepel)
  Data=Data%>%.[.$gene_symbol%in%pathway_genes,]
  # add a column of NAs
  de=Data %>%mutate(diffexpressed=ifelse(padj>0.05|is.na(padj),"NO",ifelse(log2FoldChange>0,"UP","DOWN")))

  de$delabel <- NA
 
  de$delabel[de$diffexpressed != "NO"] <- de$gene_symbol[de$diffexpressed != "NO"]
   # plot adding up all layers we have seen so far
  ggplot(data=de, aes(x=log2FoldChange, y=-log10(padj), col=diffexpressed, label=delabel)) +
    geom_point() + ggtitle(pathway_name)+
    theme_minimal() +
    geom_text_repel() +
    scale_color_manual(values=c("DOWN"="blue", "NO"= "black","UP"= "red")) +
    geom_vline(xintercept=c(-0.6, 0.6), col="red") +
    geom_hline(yintercept=-log10(0.05), col="red")
  
  
  
  
}

Valcano_plot_rnaseq_add_label=function(All_res_Data,Pathway_list,pathway_name){
  
  cols <- c("DOWN"="#26b3ff", "NO"= "grey","UP"= "#ffad73","up" = "red", "down" = "blue") 
  sizes <- c( "DOWN"=0.5,"UP"=0.5 ,"up" = 2, "down" = 2, "NO" = 0.5) 
  alphas <- c("UP" = 1, "DOWN" = 1, "NO" = 0.5)
  
  pathway_genes=Pathway_list [[pathway_name]]%>%as.character()
  
  library(ggrepel)
  Data=All_res_Data%>%dplyr::filter(!gene_symbol%in%pathway_genes)%>%mutate(diffexpressed=ifelse(padj>0.05|is.na(padj),"NO",ifelse(log2FoldChange>0,"UP","DOWN")))%>%mutate(delabel=NA)
    
  pathway_Data=All_res_Data%>%.[.$gene_symbol%in%pathway_genes,]%>%mutate(diffexpressed=ifelse(padj>0.05|is.na(padj),"NO",ifelse(log2FoldChange>0,"up","down")))%>%mutate(delabel=ifelse(diffexpressed!="NO"&(log2FoldChange>0.27|log2FoldChange<as.numeric( "-0.27")),gene_symbol,NA))
  # add a column of NAs
  de=rbind(Data,pathway_Data)
 
  MAX_y=max(-log10(de$padj))
  # de$delabel <- NA
  # 
  # de$delabel[de$diffexpressed != "NO"] <- de$gene_symbol[de$diffexpressed != "NO"]
   # plot adding up all layers we have seen so far
  ggplot(data=de, aes(x=log2FoldChange, y=-log10(padj), col=diffexpressed, label=delabel)) +
    geom_point(show.legend = F) + ggtitle(pathway_name)+
    theme_minimal() +
    geom_text_repel() +scale_color_manual(values=cols)+scale_size_manual(values=sizes)+scale_alpha_manual(values=alphas)+
    geom_vline(xintercept=c(-0.27, 0.27), col="red") +
    geom_hline(yintercept=-log10(0.05), col="red")+theme_bw()+  theme(legend.position = "none", panel.border = element_rect(colour = "black", fill = NA, size= 0.5),    
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank()) 
  
  
  
  
}


Valcano_plot_rnaseq_add_label_update=function(All_res_Data,Pathway_list,pathway_name){
  
  cols <- c("DOWN"="#26b3ff", "NO"= "grey","UP"= "#ffad73","up" = "red", "down" = "blue") 
  sizes <- c( "DOWN"=0.5,"UP"=0.5 ,"up" = 2, "down" = 2, "NO" = 0.5) 
  alphas <- c("UP" = 1, "DOWN" = 1, "NO" = 0.5)
  
  pathway_genes=Pathway_list [[pathway_name]]%>%as.character()
  
  library(ggrepel)
  Data=All_res_Data%>%dplyr::filter(!gene_symbol%in%pathway_genes)%>%mutate(diffexpressed=ifelse(padj>0.05|is.na(padj),"NO",ifelse(log2FoldChange>0,"UP","DOWN")))%>%mutate(delabel=NA)
    
  pathway_Data=All_res_Data%>%.[.$gene_symbol%in%pathway_genes,]%>%mutate(diffexpressed=ifelse(padj>0.05|is.na(padj),"NO",ifelse(log2FoldChange>0,"up","down")))%>%mutate(delabel=ifelse(diffexpressed!="NO"&(log2FoldChange>0.27|log2FoldChange<as.numeric( "-0.27")),gene_symbol,NA))
  # add a column of NAs
  de=rbind(Data,pathway_Data)
 
  MAX_y=max(-log10(de$padj))
  # de$delabel <- NA
  # 
  # de$delabel[de$diffexpressed != "NO"] <- de$gene_symbol[de$diffexpressed != "NO"]
   # plot adding up all layers we have seen so far
  ggplot(data=de, aes(x=log2FoldChange, y=-log10(padj), col=diffexpressed, label=delabel)) +
    geom_point(show.legend = F) + ggtitle(pathway_name)+
    theme_minimal() +
    geom_text_repel(size=3,force_pull= 0,max.overlaps  = Inf,
    data = subset(de, log2FoldChange >2),
    nudge_x= 6 - subset(de, log2FoldChange >2)$log2FoldChange,
    segment.size= 0.2,
    segment.color= "grey50",
    direction = "y",
    hjust         = 0.5 ) +
    
    geom_text_repel(size=3,force_pull= 0,max.overlaps  = Inf,
    data = subset(de, log2FoldChange >1&log2FoldChange<2),
    nudge_x= 4.5 - subset(de, log2FoldChange >1&log2FoldChange<2)$log2FoldChange,
    segment.size= 0.2,
    segment.color= "grey50",
    direction = "y",
    hjust         = 0.5 ) +
    
    geom_text_repel(size=3,force_pull= 0,max.overlaps  = Inf,
    data = subset(de, log2FoldChange < as.numeric(-1)),
    nudge_x= -5 - subset(de, log2FoldChange <as.numeric(-1))$log2FoldChange,
    segment.size= 0.2,
    segment.color= "grey50",
    direction = "y",
    hjust         = 0.5 ) +scale_color_manual(values=cols)+scale_size_manual(values=sizes)+scale_alpha_manual(values=alphas)+
    geom_vline(xintercept=c(-0.27, 0.27), col="red") +
    geom_hline(yintercept=-log10(0.05), col="red")+theme_bw()+  theme(legend.position = "none", panel.border = element_rect(colour = "black", fill = NA, size= 0.5),    
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank()) 
  
  
  
  
}



# Batch_boxplot_4=function(pathway_list,Res,data,x_name,fillby,nCol,Compare_list,data_type) {
#   library(gridExtra)
#   
#   plist=list()
#   for(i in 1:length(pathway_list)) {
#     Pathway=pathway_list[i]
#     # return(Pathway)
#     plist[[i]]=Bar_plot_RNAseq_4(data,Res,Pathway,x_name,fillby,Compare_list,data_type )
#     
#   }
#   #plist[["legend"]]=UMAP_legend_2(data, pathway_list[1] ,fillby,fillby)
#   
#   do.call("grid.arrange", c(plist, ncol=nCol))
#   
# }



Bar_plot_RNAseq_4=function(data,Res,pathway_name,x_name,fill_name,compare_list,data_type) {
  library(ggpubr)
  my_comparisons <- compare_list
  stat.test=Res[rownames(Res) %in% pathway_name,]
  
  Data=data[,c(pathway_name,"Group")] %>%setNames(c("Gene","Group"))
 
  yposition=max(Data$Gene)
  stat.test=Res[rownames(Res) %in% pathway_name,]  %>% mutate(y.position = 1,".y."="Gene")%>%
    mutate("p.format"=ifelse(padj<0.001,"<0.001",ifelse(padj<0.01,"<0.01",ifelse(padj<0.05,"<0.05",round(padj,2)))) ,method="Deseq2")  %>%
    rename(p=pvalue,"p.adj"="padj",)   %>% mutate(y.position = yposition+0.5 ) 
  # p=ggplot(data,aes_string(
  #   x =x_name,
  #   y = pathway_name,
  #   fill=fill_name
  # ) )+  geom_violin( ) +
  #   geom_boxplot(width=0.1) + theme(plot.title=element_text(size=5),legend.position="none")+ scale_fill_brewer(palette="Dark2")+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
  
  p=ggboxplot(Data, x = "Group", y = "Gene",fill="Group")+theme(plot.title=element_text(size=5),legend.position="none")+ylab(paste0(pathway_name,"_",data_type))+xlab("")+scale_fill_brewer(palette="Dark2")+ggtitle(pathway_name)+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +theme(plot.title = element_text(size = 10, face = "bold"))
  p + stat_pvalue_manual(stat.test,label = "p.format")
  
}
Heatmap_simple=function(gene_list,expression_data,expression_data_anno,Annotation_File){
   
  Annotation_file=Annotation_File[,c("Group")]  %>%as.data.frame()
  rownames(Annotation_file)=Annotation_File$Sample_name
  pheatmap_pathway_3_no_genename (expression_data_anno,unlist(gene_list),names(gene_list), Annotation_file,"V1")
  #list(p2)
}


Enrich_dotplot=function(FGSEA_out,top_number){
  library(ggbreak)
  FGSEA_out$NES=round(FGSEA_out$NES,1)
  #mycolor=RColorBrewer::brewer.pal(3,"YlOrRd")

if(nrow(FGSEA_out)<top_number) {
    data=FGSEA_out%>%mutate(Norm_padj=-log10(padj))%>%mutate(State=ifelse(NES>0,"Activated","Inhibited"))%>%arrange(desc(NES))
    Up_data=data%>%dplyr::filter(NES>0)%>%arrange(NES)
    DN_data=data%>%dplyr::filter(NES<0)
    data=rbind(DN_data,Up_data)
  data$pathway=factor(data$pathway,levels =data$pathway )
  MAX_NES=max(data$NES)%>%ceiling()
  min_NES=min(data$NES)%>%floor()
  # ggplot(data) +geom_point(aes(x = NES, y = pathway, color = -log10(padj),size=gene_counts))+scale_size("gene_counts") +
  P= ggplot(data) +geom_point(aes(x = NES, y = pathway, color =State ,size=Norm_padj))+scale_size("Norm_padj") +
     theme_bw() +labs(x="Normalized enrichment score",y="")+
      theme(axis.text.x = element_text(size=rel(1.15)),
        axis.title = element_text(size=rel(1.15)),plot.title = element_text(hjust=0.5, 
      	face = "bold"),legend.title = element_text(size=rel(1.15),hjust=0.5, face="bold"),axis.text.y = element_text(size =15))+scale_fill_manual(name="State",c("red","green"))+scale_size_continuous (name="-log10(padj)")
  if (min_NES>0|MAX_NES<0) {
    return(P)
    } else {
    Breaks=c(min_NES,-1,1,MAX_NES)
   return(P+scale_x_break(c(-1, 1), scales = 1.5))
 # xlab("Normalized enrichment score") +ylab("")+scale_x_break(c(-1, 1), scales = 1.5)
   }
  
  

 } else {
      data=FGSEA_out%>%arrange(padj)%>%mutate(Norm_padj=-log10(padj))%>%mutate(State=ifelse(NES>0,"Activated","Inhibited"))%>%arrange(desc(NES))
       Up_data=data%>%dplyr::filter(NES>0)%>%arrange(NES)
      DN_data=data%>%dplyr::filter(NES<0)
      if(nrow(Up_data)>nrow(DN_data)){
           DN_top_number=top_number/2
        up_top_number=top_number-DN_top_number  
       
      } else {
        up_top_number=top_number/2
        DN_top_number=top_number-up_top_number

      
      }
      
      UP_data2=Up_data%>%arrange(NES)%>%.[1:up_top_number,]%>%na.omit()
      UP_data2$pathway=factor(UP_data2$pathway,levels =UP_data2$pathway )

     DN_data2=DN_data%>%arrange(NES)%>%.[1:DN_top_number,]%>%na.omit()
      DN_data2$pathway=factor( DN_data2$pathway,levels = DN_data2$pathway )
  
      P1=ggplot(UP_data2) +geom_point(aes(x = NES, y = pathway, size=Norm_padj), color ="red")+scale_size("Norm_padj") +
     theme_bw() +labs(x="Normalized enrichment score",y="")+
      theme(axis.text.x = element_text(size=rel(1.15)),
        axis.title = element_text(size=rel(1.15)),plot.title = element_text(hjust=0.5, 
      	face = "bold"),legend.title = element_text(size=rel(1.15),hjust=0.5, face="bold"),axis.text.y = element_text(size =15))+scale_size_continuous (name="-log10(padj)") +ggtitle("Activated pathways")
    
        P2=ggplot(DN_data2) +geom_point(aes(x = NES, y = pathway ,size=Norm_padj), color ="green")+scale_size("Norm_padj") +
     theme_bw() +labs(x="Normalized enrichment score",y="")+
      theme(axis.text.x = element_text(size=rel(1.15)),
        axis.title = element_text(size=rel(1.15)),plot.title = element_text(hjust=0.5, 
      	face = "bold"),legend.title = element_text(size=rel(1.15),hjust=0.5, face="bold"),axis.text.y = element_text(size =15))+scale_size_continuous (name="-log10(padj)")+ggtitle("Inhibited pathways")
        
      cowplot::plot_grid(P1,P2,nrow=1)  
    
     }
}


# pheatmap_pathway_3_no_genename<- function (expression_data, Gene_select,pathway_name,Annotation,Label_def) {
#   Gene_select_anno= expression_data[,colnames(expression_data) %in% Gene_select] %>%t()
#   pheatmap::pheatmap(Gene_select_anno, cellwigermline= 2, cellheight =4, scale = "row",cellwidth = 4,
#                      treeheight_row = 5,main=pathway_name,
#                      show_rownames = F,show_colnames = F,
#                      annotation_col= Annotation,
#                      # annotation_row=Annotation,
#                      #annotation_legend=Label_def,
#                      cluster_rows = T,  cluster_cols = F,clustering_distance_rows = "euclidean")
#   
#   
# }
```



```{r,include=F}
# “MEISSNER_BRAIN_HCP_WITH_H3K27ME3”, “WINNEPENNINCKX_MELANOMA_METASTASIS_UP”, and “HALLMARK_MTORC1_ SIGNALING”, 
# key_words_enrichment=Pathway_enrichment_analysis(All_pathway_singscore_q005$pathway,5)
# Int_key_words=c("H3K27ME3","P53","TP53","IL4","IL12","TGFB","TNF","MESENCHYMAL","APOPTOSIS","CYCLE","METASTASIS","MITOTIC","STROMAL","EPITHELIAL","MESENCHYMAL","AKT","PI3K","PATHWAY","MYC","EGFR","ERBB","WNT","JNK","P38","RHO","ARF","PDGF","GROWTH_FACTOR","MHC")
# Key_pathway=do.call("rbind", Pathway_select_by_keyword(All_pathway_singscore_q005$pathway,Int_key_words) ) %>%as.data.frame() %>%setNames("Keyword") %>%
#        mutate(Keyword=gsub("TP53","P53" ,Keyword)) %>%arrange(Keyword)
# Key_pathway_anno=All_pathway_singscore_q005 %>%.[.$pathway%in% rownames(Key_pathway) ,] %>%.[,c("pathway","padj","NES")]

Sig_pathways=c("MEISSNER_BRAIN_HCP_WITH_H3K27ME3","HALLMARK_MTORC1_SIGNALING" ,"WINNEPENNINCKX_MELANOMA_METASTASIS_UP")
Sig_pathways_anno=All_pathway_singscore%>%dplyr::filter(pathway%in%Sig_pathways)

knitr::kable(Sig_pathways_anno[,c(1,3,6)],caption = "Significant pathways in MPNST (vs PN) " ) %>%kableExtra::kable_styling(latex_options = c("hold_position"))
```


```{r}
Disease=data.frame(tumorType=c("Plexiform Neurofibroma","Cutaneous Neurofibroma", "Diffuse Infiltrating Neurofibroma" ,"Nodular Neurofibroma" ,"Malignant Peripheral Nerve Sheath Tumor"))%>%mutate(Disease=c("PN","cNF","DifNF","NodNF","MPNST"  ))

RNAseq_anno=read.csv(file.path(COUNT_path,"RNAseq_samplemap.csv"))%>%mutate(batch=ifelse(specimenID=="JH-2-023-91HBD","WU_batch2",batch))%>%mutate(Sample_name=Syn_speicimenID)%>%left_join(.,Disease)

```

```{r}
UMAP_plot=function(Data_select,annotation) {
 Data.umap = umap::umap(Data_select)
 umap_plot_df <- data.frame(Data.umap$layout)%>%setNames(c("UMAP1","UMAP2"))
 rownames(umap_plot_df)==rownames(annotation)
 umap_plot_df_2=cbind(umap_plot_df,Data_select)%>%tibble::rownames_to_column(.,var="Sample_name")%>%left_join(.,annotation,by="Sample_name")

 
ggplot(
  umap_plot_df_2,
  aes(
    x = UMAP1,
    y = UMAP2,
    color=tumorType
  )
) +  geom_point( size = 3)

}
```


```{r}
MPNST_PN_RNAseq_anno=RNAseq_anno%>%dplyr::filter(Disease%in%c("MPNST","PN"))
Expression_vst_anno_MPNST_PN=Expression_vst_anno%>%dplyr::filter(Sample_name%in%MPNST_PN_RNAseq_anno$Sample_name)
Expression_vst_MPNST_PN=Expression_vst[,rownames(Expression_vst_anno_MPNST_PN)]
MPNST_vs_PN_res_2=MPNST_vs_PN_res%>%mutate(gene_symbol=rownames(.))
MPNST_vs_PN_res_q005=MPNST_vs_PN_res %>%dplyr::filter(padj<0.05)

q005_gene=rownames(MPNST_vs_PN_res_q005)
Expression_vst_anno_MPNST_PN_q005=Expression_vst_anno_MPNST_PN[,c("Sample_name",q005_gene)]
Expression_vst_MPNST_PN_q005=Expression_vst_MPNST_PN[q005_gene,]

MPNST_PN_sample=MPNST_PN_RNAseq_anno [,c("Sample_name","individualID","tumorType","Disease")]

```


```{r}
sig_gene=MPNST_vs_PN_res %>%dplyr::filter(padj<0.01)%>%rownames(.)
Expression_vst_MPNST_PN_sig=Expression_vst_MPNST_PN[sig_gene,]


```


\newpage


```{r,include=T}

## MPNST markers (S100B, CNP,PMP22,ngfr ,Expression of Schwann cell differentiation markers Downregulaed in MPNST, TE-cadherin (CDH1) was downregulated in both neurofibroma and MPNSTs; TWIST1,SOX9,SOX10  ,stem cell marker, upregulated in MPNST,  GAS1 (Growth Arrest Specific 1), IFG2, PTK7, FGFR1, TWIST1, and EGFR.upregulated in MPNST)

MPNST_associated_genes= c( "UBR5","TP53", "CNP","PMP22","NGFR","SOX9","SOX10","S100B","S100A4", "S100A1", "S100A2", "S100A6","CCNB2","TWIST1","EGFR","FGFR1","ERBB3","ERBB2","COL6A3","CDH1","CDKN2A","PDGFA","PDGFB","TGFBR2","GAS1","TGFB1","TGFB2","IGF2","PTK7")

Pathway_list=list("MPNST_associated_genes"=MPNST_associated_genes)

Stat_method="wilcox.test"
gene_list2=data.frame(MPNST_associated_genes=MPNST_associated_genes)
colnames(Expression_vst_MPNST_PN_q005)==MPNST_PN_sample$Sample_name




```

# Figure 1

```{r,include=T,fig.width=8,fig.height=6}
set.seed(123456)
Tumor_color=tumortype_colors%>%na.omit()

mat_colors <- list(Group = Tumor_color[1:2])
names(mat_colors$Group) <- c("PN","MPNST")

Manu_color=unlist(mat_colors)
  names(Manu_color)=gsub("Group.", "",names(Manu_color))


P1=PCA_Plot_3  (t(Expression_vst),RNAseq_anno,"Sample_name","tumorType",Tumor_color)+theme(legend.position = "none")

```


```{r,include=T,fig.width=8,fig.height=6}
P2=UMAP_plot(t(Expression_vst),RNAseq_anno)+scale_color_manual(values = Tumor_color)

```

```{r,include=T,fig.width=14,fig.height=7}
cowplot::plot_grid(P1,P2,rel_widths = c(1,1.5),labels = c("A","B"))
```



```{r,include=F,fig.height=6,fig.width=8}
#P2=Valcano_plot_rnaseq (MPNST_vs_PN_res_2,Pathway_list,"MPNST_associated_genes")
P3=Valcano_plot_rnaseq_add_label(MPNST_vs_PN_res_2,Pathway_list,"MPNST_associated_genes")
```


```{r,include=F,fig.height=6,fig.width=8}
P4=Singscore_data_plot_2(gene_list2,Expression_vst_MPNST_PN_q005,Expression_vst_anno_MPNST_PN_q005,GTAC_Sample_anno,Stat_method,mat_colors)

```




```{r,include=T,fig.width=12,fig.height=8}
#gridExtra:: grid.arrange(P1,P2, P3[[1]][[4]], nrow = 2)

#cowplot:: plot_grid(P3,NULL,P4[[1]][[4]],NULL,  nrow = 2,labels = c("C","","D"),rel_widths = c(0.5,1,1,0.1))
cowplot:: plot_grid(P3,P4[[1]][[4]],nrow = 1,labels = c("C","D"),scale = c(0.7,1))

```



```{r,eval=F}
MPNST_sig_markers=MPNST_vs_PN_res_2%>%.[MPNST_associated_genes,]%>%dplyr::filter(padj<0.05)%>%arrange(desc(log2FoldChange))

knitr::kable(MPNST_sig_markers[,c(2,6)],caption = "Significant expression change of MPNST markers in MPNST (vs PN, p<0.05, FC>1.2 or FC<-1.2) ")
```


```{r,eval=F,fig.cap="MPNST markers expression box plot",include=T,fig.height=20,fig.width=14,out.height="95%"}
Compare_list=list(c("MPNST","PN"))
setdiff(MPNST_associated_genes,colnames(Expression_vst_anno_MPNST_PN))

res_2=MPNST_vs_PN_res%>%mutate(group1="MPNST",group2="PN")
Batch_boxplot_4(MPNST_sig_markers$gene_symbol,res_2,Expression_vst_anno_MPNST_PN,x_name="Group",fillby="Group",nCol=4,Compare_list=Compare_list,"Normalized_counts")
```



```{r,include=F}
# Significant pathways

# 1361 genes with pvalue <0.05 were involved in the analysis


GMT_files=list.files(file.path(Reference_path,"2023_2"),full.names = T)%>%.[grepl("gmt.txt",.)]%>%.[grepl("c2.|h",.)]

res_P005_rank=MPNST_vs_PN_res_q005%>%arrange(log2FoldChange) %>%.[c("log2FoldChange")]%>%as.data.frame() %>%tibble::rownames_to_column()

ranks <- tibble::deframe(res_P005_rank)
fgsea_list=list()
for (i in GMT_files){
 name=gsub(".*/","",i)
 name2=gsub("symbols.gmt.txt","",name)
 pathway= fgsea::gmtPathways(i)
 fgseaRes <- fgsea::fgseaMultilevel(pathway, ranks,
                           minSize=5,
                           maxSize=500
                           )
 fgsea_list[[name2]]=fgseaRes

}

Fgsea_all=do.call("rbind",fgsea_list)
Fgsea_all_sig=Fgsea_all%>%dplyr::filter(padj<0.05)%>%arrange(padj)%>%dplyr::rename(gene_counts=size)
```
`


## Hallmark pathway
```{r,include=T,fig.width=18,fig.height=8}
Hallmark_pathway= Fgsea_all_sig%>%dplyr::filter(grepl("HALLMARK",pathway))%>%.[,c(1,3,6,7)]%>%arrange(desc(NES))

knitr::kable(Hallmark_pathway,caption = "Significant Hallmark pathways in MPNST vs PN (Adjust p value <0.05)")%>%kableExtra::kable_styling(latex_options = c("hold_position"))
P5=Enrich_dotplot(Hallmark_pathway,20)
```

## Reactome pathway
```{r,include=T}
Reactome_pathway= Fgsea_all_sig%>%dplyr::filter(grepl("REACTOME",pathway))%>%.[,c(1,3,6,7)]%>%arrange(desc(NES))

knitr::kable(Reactome_pathway,caption = "Significant Reactome pathways in MPNST vs PN (Adjust p value <0.05)")%>%kableExtra::kable_styling(latex_options = c("hold_position","scale_down"))
```



```{r,fig.width=18,fig.height=12,include=T}
P6=Enrich_dotplot(Reactome_pathway,25)
```
# Figure 2

```{r,fig.width=18,fig.height=14,include=T}
cowplot:: plot_grid(P5,P6,nrow = 2,labels = c("A","B"),rel_heights = c(1,2))

```


# Figure 3



```{r,include=T,fig.width=18,fig.height=15}
GMT_files=list.files(file.path(Reference_path,"2023_2"),full.names = T)%>%.[grepl("gmt.txt",.)]%>%.[grepl("c2.|h",.)]
pathway= c(fgsea::gmtPathways(GMT_files[1]),fgsea::gmtPathways(GMT_files[2]))

Hallmark_list=list()

for (i in Hallmark_pathway$pathway){
  Pathway_list=pathway[[i]]%>%as.list()
  names(Pathway_list)=i
  Hallmark_list[[i]]=Valcano_plot_rnaseq_add_label_update(MPNST_vs_PN_res_2,pathway,i)
  
}

cowplot::plot_grid( plotlist =Hallmark_list,ncol=3   ) 
```

# Figure 4
```{r,include=T,fig.width=20,fig.height=15}

Reactome_list=list()
Reactome_pathway_2=Reactome_pathway%>%dplyr::filter(pathway%in%c("REACTOME_DEGRADATION_OF_THE_EXTRACELLULAR_MATRIX","REACTOME_EXTRACELLULAR_MATRIX_ORGANIZATION","REACTOME_CELL_CYCLE_MITOTIC","REACTOME_EGR2_AND_SOX10_MEDIATED_INITIATION_OF_SCHWANN_CELL_MYELINATION"))


for (i in Reactome_pathway_2$pathway){
  Pathway_list=pathway[[i]]%>%as.list()
  names(Pathway_list)=i
  Reactome_list[[i]]=Valcano_plot_rnaseq_add_label_update(MPNST_vs_PN_res_2,pathway,i)
  
}

cowplot::plot_grid( plotlist =Reactome_list,ncol=3   ) 
```


# Supplymentary fig1


```{r,include=F,fig.height=6,fig.width=8}
Hallmark_heatmap_list=list()

for (i in Hallmark_pathway$pathway){
    Pathway_list=pathway[[i]]%>%as.list()
    names(Pathway_list)=i
    Pathway_list=pathway[[i]]%>%as.data.frame()%>%setNames(i)
    Stat_method="wilcox.test"
# gene_list2=GSEA_pathway_list[[Pathway]]%>%as.data.frame() %>%setNames(Pathway)

    Hallmark_heatmap_list[[i]]=Singscore_data_plot_2(Pathway_list,Expression_vst_MPNST_PN_q005,Expression_vst_anno_MPNST_PN_q005,GTAC_Sample_anno,Stat_method,mat_colors)[[1]]$gtable
}
```


```{r,include=T,fig.height=26,fig.width=22}
cowplot::plot_grid(plotlist = Hallmark_heatmap_list,ncol=3,scale= 1)

```

# Supplymentary fig2

```{r,include=F,fig.height=6,fig.width=8}
Reactome_heatmap_list=list()

for (i in Reactome_pathway_2$pathway){
    Pathway_list=pathway[[i]]%>%as.list()
    names(Pathway_list)=i
    Pathway_list=pathway[[i]]%>%as.data.frame()%>%setNames(i)
    Stat_method="wilcox.test"
# gene_list2=GSEA_pathway_list[[Pathway]]%>%as.data.frame() %>%setNames(Pathway)

 
    Reactome_heatmap_list[[i]]=Singscore_data_plot_2(Pathway_list,Expression_vst_MPNST_PN_q005,Expression_vst_anno_MPNST_PN_q005,GTAC_Sample_anno,Stat_method,mat_colors)[[1]]$gtable
}
```

```{r,include=T,fig.height=34,fig.width=22}
cowplot::plot_grid(plotlist = Reactome_heatmap_list,ncol=3,scale= 1)

```

```{r}
knitr::knit_exit()
```


## KEGG pathway
```{r,include=T,eval=F}
KEGG_pathway= Fgsea_all_sig%>%dplyr::filter(grepl("KEGG",pathway))%>%.[,c(1,3,6)]%>%arrange(desc(NES))

knitr::kable(KEGG_pathway,caption = "KEGG pathway (Adjust p value <0.05)")%>%kableExtra::kable_styling(latex_options = c("hold_position","scale_down"))
```

```{r,fig.width=10,fig.height=4,include=T}
Enrich_dotplot(KEGG_pathway,20)

```
\newpage

## Other pathway
```{r,include=F,eval=F}
Other_pathway= Fgsea_all_sig%>%dplyr::filter(!grepl("KEGG|REAC|WP_|HALL",pathway))%>%.[,c(1,3,6)]
Up_other_pathway=Other_pathway%>%dplyr::filter(NES>0)%>%arrange(desc(NES))
DN_other_pathway=Other_pathway%>%dplyr::filter(NES<0)%>%arrange(NES)
knitr::kable(Up_other_pathway,caption = "Upregulated pathways in Other database (Adjust p value <0.05)",longtable = TRUE)%>%kableExtra::kable_styling(latex_options = c("hold_position","repeat_header","scale_down"))

knitr::kable(DN_other_pathway,caption = "Downregulated pathways in Other database (Adjust p value <0.05)")%>%kableExtra::kable_styling(latex_options = c("hold_position","scale_down"))
```

\newpage

```{r,fig.width=16,fig.height=8,include=T,eval=F}
Enrich_dotplot(Other_pathway,20)

```
```{r}
knitr::knit_exit()
```



```{r,include=F}
# Download gene list from https://www.gsea-msigdb.org/gsea/msigdb/cards/MEISSNER_BRAIN_HCP_WITH_H3K27ME3;HALLMARK_MTORC1_SIGNALING.v2023.2.Hs.gmt;WINNEPENNINCKX_MELANOMA_METASTASIS_UP.v2023.2.Hs.gmt

Three_pathways=c("MEISSNER_BRAIN_HCP_WITH_H3K27ME3.v2023.2.Hs.gmt","WINNEPENNINCKX_MELANOMA_METASTASIS_UP.v2023.2.Hs.gmt","HALLMARK_MTORC1_SIGNALING.v2023.2.Hs.gmt")
plist=list()
for (i in Three_pathways){
  Pathway=gsub(".v2023.2.Hs.gmt","",i)
 Stat_method="wilcox.test"
 pathway=file.path(Reference_path,i)
 genelist= fgsea::gmtPathways(pathway)
 P2=Valcano_plot_rnaseq (MPNST_vs_PN_res_2,genelist,Pathway)
 P3=Singscore_data_plot_2(genelist,Expression_vst_MPNST_PN_q005,Expression_vst_anno_MPNST_PN_q005,GTAC_Sample_anno,Stat_method,mat_colors)
 plist[[i]]=cowplot:: plot_grid(P2, P3[[2]], align = "v", ncol=2, scale = c( 1,  1))
}

```


```{r,include=T,fig.width=10,fig.height=14}
cowplot:: plot_grid(plotlist = plist, align = "v", ncol=1,labels = "AUTO", scale = c( 1,  1))
```


```{r,include=T,fig.height=16,fig.width=12}
plist2=list()
for (i in Three_pathways){
  Pathway=gsub(".v2023.2.Hs.gmt","",i)
 Stat_method="wilcox.test"
 pathway=file.path(Reference_path,i)
 genelist= fgsea::gmtPathways(pathway)
 P3=Singscore_data_plot_2(genelist,Expression_vst_MPNST_PN_q005,Expression_vst_anno_MPNST_PN_q005,GTAC_Sample_anno,Stat_method,mat_colors)
 plist2[[i]]=P3[[1]][[4]]
}
```




```{r,,fig.height=16,fig.width=10,include=T}
#cowplot:: plot_grid(plotlist = plist2,ncol=1,align = "v", labels = "AUTO", scale = c( 1,  1,1))

```

```{r}
knitr::knit_exit()
```


\newpage

```{r,fig.cap="MEISSNER_BRAIN_HCP_WITH_H3K27ME3_Heatmap_with_name",fig.height=16,fig.width=8,include=T}
# Heatmap with name
Output=Singscore_data_plot_2(genelist,Expression_vst_MPNST_PN_q005,Expression_vst_anno_MPNST_PN_q005,GTAC_Sample_anno,Stat_method,mat_colors)

```

\newpage

```{r,fig.cap="MEISSNER_BRAIN_HCP_WITH_H3K27ME3_singscore",include=T,fig.width=8,fig.height=6}
Output[[1]]
```


```{r,fig.cap="MEISSNER_BRAIN_HCP_WITH_H3K27ME3_volcano_plot",include=T,fig.width=8,fig.height=6}
Valcano_plot_rnaseq (MPNST_vs_PN_res_2,GSEA_pathway_list,Pathway)

```

```{r}
knitr::knit_exit()
```


## FISCHER G2 M CELL CYCLE

```{r,fig.height=8,fig.width=6,include=F}
Pathway="FISCHER_G2_M_CELL_CYCLE" 
Stat_method="wilcox.test"
gene_list2=GSEA_pathway_list[[Pathway]]%>%as.data.frame() %>%setNames(Pathway)
Heatmap_simple(gene_list2,Expression_vst_MPNST_PN_q005,Expression_vst_anno_MPNST_PN_q005,GTAC_Sample_anno)

```


```{r,fig.height=12,fig.width=8,include=T}
Output=Singscore_data_plot_2(gene_list2,Expression_vst_MPNST_PN_q005,Expression_vst_anno_MPNST_PN_q005,GTAC_Sample_anno,Stat_method)
```

\newpage

```{r,include=T,fig.width=8,fig.height=6}
Output[[1]]
Valcano_plot_rnaseq (MPNST_vs_PN_res_2,GSEA_pathway_list,Pathway)

```

## HALLMARK EPITHELIAL MESENCHYMAL TRANSITION

```{r,fig.height=8,fig.width=6}
Pathway="HALLMARK_EPITHELIAL_MESENCHYMAL_TRANSITION" 
Stat_method="wilcox.test"
gene_list2=GSEA_pathway_list[[Pathway]]%>%as.data.frame() %>%setNames(Pathway)
Heatmap_simple(gene_list2,Expression_vst_MPNST_PN_q005,Expression_vst_anno_MPNST_PN_q005,GTAC_Sample_anno)

```


```{r,fig.height=12,fig.width=8,include=T}
Output=Singscore_data_plot_2(gene_list2,Expression_vst_MPNST_PN_q005,Expression_vst_anno_MPNST_PN_q005,GTAC_Sample_anno,Stat_method)
```

\newpage

```{r,include=T,fig.width=8,fig.height=6}
Output[[1]]
Valcano_plot_rnaseq (MPNST_vs_PN_res_2,GSEA_pathway_list,Pathway)

```

## WINNEPENNINCKX MELANOMA METASTASIS UP


```{r,fig.height=8,fig.width=6}
Pathway="WINNEPENNINCKX_MELANOMA_METASTASIS_UP" 
Stat_method="wilcox.test"
gene_list2=GSEA_pathway_list[[Pathway]]%>%as.data.frame() %>%setNames(Pathway)
Heatmap_simple(gene_list2,Expression_vst_MPNST_PN_q005,Expression_vst_anno_MPNST_PN_q005,GTAC_Sample_anno)

```


```{r,fig.height=12,fig.width=8}
Output=Singscore_data_plot_2(gene_list2,Expression_vst_MPNST_PN_q005,Expression_vst_anno_MPNST_PN_q005,GTAC_Sample_anno,Stat_method)

```


```{r}
Output[[1]]
Valcano_plot_rnaseq (MPNST_vs_PN_res_2,GSEA_pathway_list,Pathway)

```

```{r}
knitr::knit_exit()
```


