---
title: "RNAseq Figures output for paper"
author: ""
date: "`r Sys.Date()`"
output:
  pdf_document:
    keep_tex: false
    latex_engine: xelatex
    fig_caption: yes
    highlight: haddock
      
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo =F,message = F,warning = F,include = F)
```


```{r,include=F}
# setting up working path 
rm(list=ls())
Path="/Volumes/lyu.yang/MAC_1/R/Project/11_JH_biobank_paper"
COUNT_path=file.path(Path,"data" ,"RNAseq","Count")
dir(COUNT_path)
Unused_samples=c("JH-2-084","JH-2-106") # These samples are not inclucded in the analyssi
Reference_path="/Volumes/lyu.yang/MAC_1/R/Public_database/Reference/Human_GSEA" # Download GMT files from  from https://www.gsea-msigdb.org/gsea/msigdb
Process_data=file.path(Path,"data","Process_data")

dir.create(COUNT_path,recursive = T)
dir(Reference_path)
```


```{r,include=F}
library(dplyr)
library(ggplot2)
library(ggbreak)
library(synapser)
 
```


```{r}
# Functions
PCA_Plot_3<- function (data,Annotation,VAR,Color,manu_color) {
  # logcountdata row:genes,column: samples
  pca <- prcomp(data) 
  pca_out<-as.data.frame(pca$x)
  df_out<- pca_out %>%tibble::rownames_to_column(var=VAR)  %>% left_join(., Annotation) 
  #df_out<- merge (pca_out,Annotation,by.x=0,by.y=0) 
  PC1_var=summary(pca)$importance%>%as.data.frame()%>%.[2,1]%>%scales:: percent(accuracy = 0.01)
  PC2_var=summary(pca)$importance%>%as.data.frame()%>%.[2,2]%>%scales:: percent(accuracy = 0.01)
  # label_color<- factor(df_out[,group])
  ggplot(df_out,aes_string(x="PC1",y="PC2")) +geom_point(aes_string(colour = Color),size=3) +scale_color_manual(values = manu_color)+xlab(paste0("PC1(",PC1_var,")"))+ylab(paste0("PC2(",PC2_var,")"))+labs(color = gsub("_"," ",Color))+
    theme(legend.position = "bottom",legend.direction = "vertical",legend.title = element_text(size=12, vjust = .5, hjust = .3))
}



pheatmap_pathway_3<- function (expression_data, Gene_select,pathway_name,Annotation,Label_def,anno_color) {
 # Gene_select<-Pathway_list[,pathway_name]%>%unique()
  Gene_select_anno= expression_data[,colnames(expression_data) %in% Gene_select] %>%t()
  pheatmap::pheatmap(Gene_select_anno, cellheight =10,cellwidth =10  , scale = "row",
                     treeheight_row = 5,main=pathway_name,
                     show_rownames = T,show_colnames = F,
                     annotation_col= Annotation,
                     annotation_colors= anno_color,border_color = "black",
                     # annotation_row=Annotation,
                     #annotation_legend=Label_def,
                     cluster_rows = T,  cluster_cols = F,clustering_distance_rows = "euclidean",silent = F)
  
}


pheatmap_pathway_3_update <- function(expression_data, Gene_select, pathway_name,
                               Annotation, legend_title, anno_color) {
  # Subset the expression matrix for the selected genes and transpose it
  Gene_select_anno <- t(expression_data[, colnames(expression_data) %in% Gene_select])
  
  # Create the heatmap using pheatmap with silent = TRUE to capture the gtable object
  ph <- pheatmap::pheatmap(Gene_select_anno,
                           cellheight = 10,
                           cellwidth  = 10,
                           scale = "row",
                           treeheight_row = 5,
                           main = pathway_name,
                           show_rownames = TRUE,
                           show_colnames = FALSE,
                           annotation_col = Annotation,
                           annotation_colors = anno_color,
                           border_color = "black",
                           cluster_rows = TRUE,
                           cluster_cols = FALSE,
                           clustering_distance_rows = "euclidean",
                           silent = TRUE)
  
  # Load required packages
  require(grid)
  require(gtable)
  
  # Extract the gtable object from the pheatmap output
  g <- ph$gtable
  
  # Locate the legend in the gtable layout (its name is "legend")
  legend_index <- which(g$layout$name == "legend")
  
  if (length(legend_index) > 0) {
    # Extract the legend grob
    legend <- g$grobs[[legend_index]]
    
    # If the legend is not already a gtable, wrap it in a gtable.
    if (!inherits(legend, "gtable")) {
      legend <- gtable::gtable(widths = unit(1, "npc"),
                               heights = unit(1, "npc"),
                               grobs = list(legend))
    }
    
    # Remove the original legend from its default location
    g$grobs[[legend_index]] <- grid::nullGrob()
    
    # Create a text grob for the legend title
    title_grob <- grid::textGrob(legend_title, 
                                 gp = grid::gpar(fontsize = 10, fontface = "bold"))
    
    # Add a new row at the top of the legend for the title, with a small extra height for spacing
    legend <- gtable::gtable_add_rows(legend,
                                      heights = grid::grobHeight(title_grob) + unit(2, "mm"),
                                      pos = 0)
    legend <- gtable::gtable_add_grob(legend,
                                      title_grob,
                                      t = 1, l = 1, r = ncol(legend))
    
    # Add a new row at the bottom of the overall gtable to accommodate the legend
    g <- gtable::gtable_add_rows(g,
                                 heights = grid::grobHeight(legend) + unit(4, "mm"))
    # Place the modified legend into this new row spanning the full width
    g <- gtable::gtable_add_grob(g, legend,
                                 t = nrow(g), l = 1, r = ncol(g))
  }
  
  # Draw the modified heatmap
  grid::grid.newpage()
  grid::grid.draw(g)
  
  invisible(ph)
}


singcore_Normalized_GSEA=function (data,genelist, is.na=T) {
  nm=names(genelist)
  plist=list()
  Data_prepare=singscore::rankGenes (data)
 # n_genes=length(intersect(genelist,rownames(data)))
  
   if( is.list(genelist)) {
    for (i in 1:length(genelist)) { 
      n_genes=length(intersect(unique(genelist[[i]]),rownames(data)))
      if (n_genes>4){
      scoredf<- singscore::simpleScore (Data_prepare, upSet=unique(genelist[[i]]))%>%.[1]
      scoredf$Normalized=scale(scoredf[,1])
      colnames(scoredf)=c(paste0(nm[i],"_singscore") , paste0(nm[i],"_Normalized_singscore" ))
      if(nrow(scoredf)==ncol(data)) {
        plist[[nm[i]]]=scoredf
      }
      }
     }
  }
  if(is.data.frame(genelist))  {
    
  for (i in seq_along(genelist)) { 
    n_genes=length(intersect(unique(genelist[[i]]),rownames(data)))
    if (n_genes>4){
    scoredf<- singscore::simpleScore (Data_prepare, upSet=unique(genelist[,i]))%>%.[1]
    scoredf$Normalized=scale(scoredf[,1])
    colnames(scoredf)=c(paste0(nm[i],"_singscore") , paste0(nm[i],"_Normalized_singscore" ))
    if(nrow(scoredf)==ncol(data)) {
      plist[[nm[i]]]=scoredf
     }
    }
  }
  }
  return(plist)
  #output=do.call("cbind",plist)
}

Singscore_data_plot_2=function(gene_list,expression_data,expression_data_anno,Annotation_File,stat_method,Anno_color){
  Gene_singscore= singcore_Normalized_GSEA(expression_data,gene_list)
  Manu_color=unlist(Anno_color )
  names(Manu_color)=gsub("Group.", "",names(Manu_color))

  Annotation_file=Annotation_File[,c("Group")]  %>%as.data.frame()
  rownames(Annotation_file)=Annotation_File$Sample_name
  plist=list()
 
  plist[[1]]=pheatmap_pathway_3 (expression_data_anno,unlist(gene_list),names(gene_list), Annotation_file,"V1",Anno_color)

  
  plist[[2]]=GGplot_singscore_boxplot_2(Gene_singscore[[1]],Annotation_File,"Group","Sample_name",Method=stat_method,Manu_color)

}


GGplot_singscore_boxplot_2=function(singscore,Annotation,group_col,Var,Method,manu_color) {
  library(ggplot2)
  library(ggpubr)
  gene_name=colnames(singscore)[2]
  Data=singscore%>%tibble::rownames_to_column(.,var=Var)
  Data2= Data %>% left_join(.,Annotation,by=Var)%>%as.data.frame()


  ggplot(Data2,aes_string(x="Group",y=gene_name,fill=group_col)) +geom_boxplot() +xlab("")+
    ylab(gene_name)+scale_fill_manual(values = manu_color)+ theme(axis.title.y = element_text(size=8),axis.text.x = element_text(angle=90, vjust=.5, hjust=1)) +stat_compare_means(method = Method)



}




Singscore_boxplot <- function(gene_list, expression_data, Annotation_File,Group, stat_method, Manu_color) {
  library(ggplot2)
  library(ggpubr)
  library(dplyr)

  # Calculate Singscore
  Gene_singscore <- singcore_Normalized_GSEA(expression_data, gene_list)

  # Get first singscore result
  singscore_data <- Gene_singscore[[1]]
  gene_name <- colnames(singscore_data)[2]  # Extract gene column name

  # Prepare Data for Boxplot
  Data <- singscore_data %>%
    tibble::rownames_to_column(var = "Sample_name") %>%
    left_join(Annotation_File, by = "Sample_name") %>%
    as.data.frame()
 
 boxplot_plot <- ggplot(Data, aes_string(x = Group, y = gene_name, fill = Group)) +
    geom_boxplot(outlier.shape = NA, alpha = 0.8) +  # Adjust transparency for clarity
    xlab("") + ylab(gsub("_"," ",gene_name)) +
    scale_fill_manual(values = Manu_color) +
    stat_compare_means(method = stat_method,size =8) +
    theme_classic() +  # Remove background
    theme(
      axis.title.y = element_text(size = 20),
      axis.text.x = element_text(size=20,angle =45, vjust = 0.7, hjust =0.7 ),
      panel.grid.major = element_blank(),  # Remove major grid lines
      panel.grid.minor = element_blank(),  # Remove minor grid lines
      panel.border = element_blank(),      # Remove panel border
      legend.position = "none"             # Remove legend
    )

  return(boxplot_plot)
}


Volcano_plot_rnaseq_add_label=function(All_res_Data,Pathway_list,pathway_name){
  
  cols <- c("DOWN"="#26b3ff", "NO"= "grey","UP"= "#ffad73","up" = "red", "down" = "blue") 
  sizes <- c( "DOWN"=0.5,"UP"=0.5 ,"up" = 2, "down" = 2, "NO" = 0.5) 
  alphas <- c("UP" = 1, "DOWN" = 1, "NO" = 0.5)
  
  pathway_genes=Pathway_list [[pathway_name]]%>%as.character()
  
  library(ggrepel)
  Data=All_res_Data%>%dplyr::filter(!gene_symbol%in%pathway_genes)%>%mutate(diffexpressed=ifelse(padj>0.05|is.na(padj),"NO",ifelse(log2FoldChange>0,"UP","DOWN")))%>%mutate(delabel=NA)
    
  pathway_Data=All_res_Data%>%.[.$gene_symbol%in%pathway_genes,]%>%mutate(diffexpressed=ifelse(padj>0.05|is.na(padj),"NO",ifelse(log2FoldChange>0,"up","down")))%>%mutate(delabel=ifelse(diffexpressed!="NO"&(log2FoldChange>0.27|log2FoldChange<as.numeric( "-0.27")),gene_symbol,NA))
  # add a column of NAs
  de=rbind(Data,pathway_Data)
 
  MAX_y=max(-log10(de$padj))
  # de$delabel <- NA
  # 
  # de$delabel[de$diffexpressed != "NO"] <- de$gene_symbol[de$diffexpressed != "NO"]
   # plot adding up all layers we have seen so far
  ggplot(data=de, aes(x=log2FoldChange, y=-log10(padj), col=diffexpressed, label=delabel)) +
    geom_point(show.legend = F) + ggtitle(pathway_name)+
    theme_minimal() +
    geom_text_repel() +scale_color_manual(values=cols)+scale_size_manual(values=sizes)+scale_alpha_manual(values=alphas)+
    geom_vline(xintercept=c(-0.27, 0.27), col="red") +
    geom_hline(yintercept=-log10(0.05), col="red")+theme_bw()+  theme(legend.position = "none", panel.border = element_rect(colour = "black", fill = NA, size= 0.5),    
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank()) 
  
  
  
  
}


Volcano_plot_rnaseq_add_label_update=function(All_res_Data,Pathway_list,pathway_name){
  
  cols <- c("DOWN"="#26b3ff", "NO"= "grey","UP"= "#ffad73","up" = "red", "down" = "blue") 
  sizes <- c( "DOWN"=0.5,"UP"=0.5 ,"up" = 2, "down" = 2, "NO" = 0.5) 
  alphas <- c("UP" = 1, "DOWN" = 1, "NO" = 0.5)
  
  pathway_genes=Pathway_list [[pathway_name]]%>%as.character()
  
  library(ggrepel)
  Data=All_res_Data%>%dplyr::filter(!gene_symbol%in%pathway_genes)%>%mutate(diffexpressed=ifelse(padj>0.05|is.na(padj),"NO",ifelse(log2FoldChange>0,"UP","DOWN")))%>%mutate(delabel=NA)
    
  pathway_Data=All_res_Data%>%.[.$gene_symbol%in%pathway_genes,]%>%mutate(diffexpressed=ifelse(padj>0.05|is.na(padj),"NO",ifelse(log2FoldChange>0,"up","down")))%>%mutate(delabel=ifelse(diffexpressed!="NO"&(log2FoldChange>0.27|log2FoldChange<as.numeric( "-0.27")),gene_symbol,NA))
  # add a column of NAs
  de=rbind(Data,pathway_Data)
 
  MAX_y=max(-log10(de$padj))
  # de$delabel <- NA
  # 
  # de$delabel[de$diffexpressed != "NO"] <- de$gene_symbol[de$diffexpressed != "NO"]
   # plot adding up all layers we have seen so far
  ggplot(data=de, aes(x=log2FoldChange, y=-log10(padj), col=diffexpressed, label=delabel)) +
    geom_point(show.legend = F) + ggtitle(pathway_name)+
    theme_minimal() +
    geom_text_repel(size=3,force_pull= 0,max.overlaps  = Inf,
    data = subset(de, log2FoldChange >2),
    nudge_x= 6 - subset(de, log2FoldChange >2)$log2FoldChange,
    segment.size= 0.2,
    segment.color= "grey50",
    direction = "y",
    hjust         = 0.5 ) +
    
    geom_text_repel(size=3,force_pull= 0,max.overlaps  = Inf,
    data = subset(de, log2FoldChange >1&log2FoldChange<2),
    nudge_x= 4.5 - subset(de, log2FoldChange >1&log2FoldChange<2)$log2FoldChange,
    segment.size= 0.2,
    segment.color= "grey50",
    direction = "y",
    hjust         = 0.5 ) +
    
    geom_text_repel(size=3,force_pull= 0,max.overlaps  = Inf,
    data = subset(de, log2FoldChange < as.numeric(-1)),
    nudge_x= -5 - subset(de, log2FoldChange <as.numeric(-1))$log2FoldChange,
    segment.size= 0.2,
    segment.color= "grey50",
    direction = "y",
    hjust         = 0.5 ) +scale_color_manual(values=cols)+scale_size_manual(values=sizes)+scale_alpha_manual(values=alphas)+
    geom_vline(xintercept=c(-0.27, 0.27), col="red") +
    geom_hline(yintercept=-log10(0.05), col="red")+theme_bw()+  theme(legend.position = "none", panel.border = element_rect(colour = "black", fill = NA, size= 0.5),    
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank()) 
  
  
  
  
}

Volcano_plot_rnaseq_add_label_update_2 <- function(All_res_Data, Pathway_list, pathway_name, top_gene_number = 10,New_title=NULL) {
  library(ggplot2)
  library(ggrepel)
  library(dplyr)
  
  if (is.null(New_title)){
    New_title=paste0(pathway_name, " Volcano Plot (Top ", top_gene_number, " Genes)")
  }

  # Define colors, sizes, and alpha values for differentially expressed genes
  cols <- c("DOWN" = "#26b3ff", "NO" = "grey", "UP" = "#ffad73", "up" = "red", "down" = "blue")
  sizes <- c("DOWN" = 0.5, "UP" = 0.5, "up" = 2, "down" = 2, "NO" = 0.5)
  alphas <- c("UP" = 1, "DOWN" = 1, "NO" = 0.5)

  # Get pathway genes
  pathway_genes <- as.character(Pathway_list[[pathway_name]])

  # Filter genes NOT in the pathway
  Data <- All_res_Data %>%
    filter(!gene_symbol %in% pathway_genes) %>%
    mutate(
      diffexpressed = case_when(
        padj > 0.05 | is.na(padj) ~ "NO",
        log2FoldChange > 0 ~ "UP",
        TRUE ~ "DOWN"
      ),
      delabel = NA
    )

  # Filter pathway genes
  pathway_Data <- All_res_Data %>%
    filter(gene_symbol %in% pathway_genes) %>%
    mutate(
      diffexpressed = case_when(
        padj > 0.05 | is.na(padj) ~ "NO",
        log2FoldChange > 0 ~ "up",
        TRUE ~ "down"
      ),
      delabel = ifelse(
        diffexpressed != "NO" & (log2FoldChange > 0.27 | log2FoldChange < -0.27),
        gene_symbol, NA
      )
    )

  # Combine datasets
  de <- bind_rows(Data, pathway_Data)

  # Get top N upregulated and downregulated genes
  top_genes_up <- pathway_Data %>% filter(diffexpressed %in% c("UP", "up")) %>% slice_max(order_by = log2FoldChange, n = top_gene_number)
  top_genes_down <- pathway_Data %>% filter(diffexpressed %in% c("DOWN", "down")) %>% slice_min(order_by = log2FoldChange, n = top_gene_number)

  # Maximum Y-axis value for scaling
  MAX_y <- max(-log10(de$padj), na.rm = TRUE)

  # Plot Volcano
  ggplot(data = de, aes(x = log2FoldChange, y = -log10(padj), color = diffexpressed, label = delabel)) +
    geom_point(size = 1.5) +  # Show points with color mapping
    scale_color_manual(values = cols) +  # Apply manual color scale
    scale_size_manual(values = sizes) + 
    scale_alpha_manual(values = alphas) + 
    theme_minimal() +
    
    # Label **Top N Upregulated Genes in Red**
    geom_text_repel(
      data = top_genes_up, 
      aes(label = gene_symbol),
      color = "red",  # 🔴 Upregulated genes labeled in red
      size = 8, segment.size = 0.5, segment.color = "grey50",
      nudge_x = 2, direction = "y", hjust = 0.5
    ) +

    # Label **Top N Downregulated Genes in Blue**
    geom_text_repel(
      data = top_genes_down, 
      aes(label = gene_symbol),
      color = "blue",  # 🔵 Downregulated genes labeled in blue
      size = 8, segment.size = 0.5, segment.color = "grey50",
      nudge_x = -3, direction = "y", hjust = 1
    ) +
    
    # Add vertical significance threshold lines
    geom_vline(xintercept = c(-0.27, 0.27), col = "red", linetype = "dashed") +
    
    # Add horizontal significance threshold line
    geom_hline(yintercept = -log10(0.05), col = "red", linetype = "dashed") +
    
    ggtitle(New_title) +
    
    # Custom theme adjustments
    theme_bw() +
    theme(
      legend.position = "none", 
      axis.text.x = element_text(size=20),
      axis.text.y = element_text(size=20),
      axis.title = element_text(size=40),
      plot.title = element_text(hjust=0.5, face = "bold",size=25),
      panel.border = element_rect(colour = "black", fill = NA, size =2),
      panel.grid.minor = element_blank(),
      panel.grid.major = element_blank()
    )
}


Bar_plot_RNAseq_4=function(data,Res,pathway_name,x_name,fill_name,compare_list,data_type) {
  library(ggpubr)
  my_comparisons <- compare_list
  stat.test=Res[rownames(Res) %in% pathway_name,]
  
  Data=data[,c(pathway_name,"Group")] %>%setNames(c("Gene","Group"))
 
  yposition=max(Data$Gene)
  stat.test=Res[rownames(Res) %in% pathway_name,]  %>% mutate(y.position = 1,".y."="Gene")%>%
    mutate("p.format"=ifelse(padj<0.001,"<0.001",ifelse(padj<0.01,"<0.01",ifelse(padj<0.05,"<0.05",round(padj,2)))) ,method="Deseq2")  %>%
    rename(p=pvalue,"p.adj"="padj",)   %>% mutate(y.position = yposition+0.5 ) 
  # p=ggplot(data,aes_string(
  #   x =x_name,
  #   y = pathway_name,
  #   fill=fill_name
  # ) )+  geom_violin( ) +
  #   geom_boxplot(width=0.1) + theme(plot.title=element_text(size=5),legend.position="none")+ scale_fill_brewer(palette="Dark2")+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
  
  p=ggboxplot(Data, x = "Group", y = "Gene",fill="Group")+theme(plot.title=element_text(size=5),legend.position="none")+ylab(paste0(pathway_name,"_",data_type))+xlab("")+scale_fill_brewer(palette="Dark2")+ggtitle(pathway_name)+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +theme(plot.title = element_text(size = 10, face = "bold"))
  p + stat_pvalue_manual(stat.test,label = "p.format")
  
}
Heatmap_simple=function(gene_list,expression_data,expression_data_anno,Annotation_File){
   
  Annotation_file=Annotation_File[,c("Group")]  %>%as.data.frame()
  rownames(Annotation_file)=Annotation_File$Sample_name
  pheatmap_pathway_3_no_genename (expression_data_anno,unlist(gene_list),names(gene_list), Annotation_file,"V1")
  #list(p2)
}


Enrich_dotplot=function(FGSEA_out,top_number=11,split=F,size_max=20,text_y_size=30){
  # set top_number > input rows, the figure will be split in up and down pathways
  library(ggbreak)
  FGSEA_out$NES=round(FGSEA_out$NES,5)
  #mycolor=RColorBrewer::brewer.pal(3,"YlOrRd")

if(nrow(FGSEA_out)<top_number) {
    data=FGSEA_out%>%mutate(Norm_padj=-log10(padj))%>%mutate(State=ifelse(NES>0,"Activated","Inhibited"))%>%arrange(desc(NES))
    Up_data=data%>%dplyr::filter(NES>0)%>%arrange(NES)
    DN_data=data%>%dplyr::filter(NES<0)
    data=rbind(DN_data,Up_data)
    data$pathway=factor(data$pathway,levels =data$pathway )
    MAX_NES=max(data$NES)%>%ceiling()
    min_NES=min(data$NES)%>%floor()
   
  # ggplot(data) +geom_point(aes(x = NES, y = pathway, color = -log10(padj),size=gene_counts))+scale_size("gene_counts") +
  #P= ggplot(data) +geom_point(aes(x = NES, y = pathway, color =State ,size=Norm_padj))+scale_size_continuous (name="-log10(padj)",range=c(2,6))+scale_fill_manual(name="State",c("red","green")) +
  #P= ggplot(data) +geom_point(aes(x = NES, y = pathway, color =State ,size=Norm_padj))+scale_size_area (name="-log10(padj)",max_size=size_max,breaks=c(2,4,6,8),limits = c(2, 10))+scale_fill_manual(name="State",c("red","green")) +
  #P= ggplot(data) +geom_point(aes(x = NES, y = pathway, color =State ,size=Norm_padj))+scale_size_continuous (name="-log10(padj)",range=c(2,size_max), breaks = c(2,  4, 6,8,10,12)),guide="legend")+scale_fill_manual(name="State",c("red","green")) +
  P= ggplot(data) +geom_point(aes(x = NES, y = pathway, color =State ,size=Norm_padj))+scale_size_area (name="-log10(padj)",max_size=size_max,breaks=c(2,4,6,8))+scale_fill_manual(name="State",c("red","green")) +  
     theme_bw() +labs(x="Normalized enrichment score",y="")+
      theme(panel.border = element_rect(colour = "black", fill = NA, size = 5),
            axis.text.x = element_text(size=20),
            axis.title = element_text(size=30),
            plot.title = element_text(hjust=0.5, face = "bold",size=15),
            legend.position = "bottom", 
            legend.direction = "vertical",
            legend.title = element_text(size=25,hjust=0.5, face="bold"),
            axis.text.y = element_text(size =text_y_size,margin = margin(r = 20)))
  if (!isTRUE(split)|min_NES>0|MAX_NES<0) {
    return(P)
    } else {
    Breaks=c(min_NES,-1,1,MAX_NES)
   return(P+scale_x_break(c(-1, 1), scales = 1.5))
 # xlab("Normalized enrichment score") +ylab("")+scale_x_break(c(-1, 1), scales = 1.5)
   }
  
  

 } else {
      data=FGSEA_out%>%arrange(padj)%>%mutate(Norm_padj=-log10(padj))%>%mutate(State=ifelse(NES>0,"Activated","Inhibited"))%>%arrange(desc(NES))
       Up_data=data%>%dplyr::filter(NES>0)%>%arrange(NES)
       DN_data=data%>%dplyr::filter(NES<0)
      if(nrow(Up_data)>nrow(DN_data)){
           DN_top_number=top_number/2
        up_top_number=top_number-DN_top_number  
       
      } else {
        up_top_number=top_number/2
        DN_top_number=top_number-up_top_number

      
      }
      
      UP_data2=Up_data%>%arrange(NES)%>%.[1:up_top_number,]%>%na.omit()
      UP_data2$pathway=factor(UP_data2$pathway,levels =UP_data2$pathway )

     DN_data2=DN_data%>%arrange(NES)%>%.[1:DN_top_number,]%>%na.omit()
      DN_data2$pathway=factor( DN_data2$pathway,levels = DN_data2$pathway )
  
      P1=ggplot(UP_data2) +geom_point(aes(x = NES, y = pathway, size=Norm_padj), color ="red")+scale_size("Norm_padj") +
     theme_bw() +labs(x="Normalized enrichment score",y="")+
      theme(axis.text.x = element_text(size=rel(1.15)),
        axis.title = element_text(size=rel(1.15)),plot.title = element_text(hjust=0.5, 
      	face = "bold"),legend.title = element_text(size=rel(1.15),hjust=0.5, face="bold"),axis.text.y = element_text(size =15))+
        scale_size_continuous (name="-log10(padj)",range=c(3,8)) +ggtitle("Activated pathways")
    
        P2=ggplot(DN_data2) +geom_point(aes(x = NES, y = pathway ,size=Norm_padj), color ="green")+scale_size("Norm_padj") +
     theme_bw() +labs(x="Normalized enrichment score",y="")+
      theme(axis.text.x = element_text(size=rel(1.15)),
        axis.title = element_text(size=rel(1.15)),plot.title = element_text(hjust=0.5, 
      	face = "bold"),legend.title = element_text(size=rel(1.15),hjust=0.5, face="bold"),axis.text.y = element_text(size =15))+scale_size_continuous (name="-log10(padj)",range=c(3,8))+ggtitle("Inhibited pathways")
        
      cowplot::plot_grid(P1,P2,nrow=1)  
    
     }
}

UMAP_plot=function(Data_select,annotation) {
 Data.umap = umap::umap(Data_select)
 umap_plot_df <- data.frame(Data.umap$layout)%>%setNames(c("UMAP1","UMAP2"))
 rownames(umap_plot_df)==rownames(annotation)
 umap_plot_df_2=cbind(umap_plot_df,Data_select)%>%tibble::rownames_to_column(.,var="Sample_name")%>%left_join(.,annotation,by="Sample_name")

 
ggplot(
  umap_plot_df_2,
  aes(
    x = UMAP1,
    y = UMAP2,
    color=tumorType
  )
) +  geom_point( size = 3)

}



pheatmap_pathway_annotation_bottom <- function(expression_data, Gene_select, pathway_name,
                                                 Annotation, anno_color) {
  # Subset the expression matrix for the selected genes and transpose it
  Gene_select_anno <- t(expression_data[, colnames(expression_data) %in% Gene_select])
  
  # Create the heatmap using pheatmap with silent = TRUE to capture the gtable object
  ph <- pheatmap::pheatmap(Gene_select_anno,
                           cellheight = 10,
                           cellwidth  = 10,
                           scale = "row",
                           treeheight_row = 5,
                           main = pathway_name,
                           show_rownames = TRUE,
                           show_colnames = FALSE,
                           annotation_col = Annotation,
                           annotation_colors = anno_color,
                           annotation_legend = T,
                           annotation_names_col=F,
                           border_color = "black",
                           cluster_rows = TRUE,
                           cluster_cols = FALSE,
                           clustering_distance_rows = "euclidean",
                           silent = TRUE)
  
  # Load required libraries
  require(grid)
  require(gtable)
  
  # Extract the gtable from the pheatmap object
  g <- ph$gtable
  
   ## 1. Modify the MAIN legend (the color scale legend) by adding a title.
  main_legend_index <- which(g$layout$name == "legend")
  if (length(main_legend_index) > 0) {
    main_legend <- g$grobs[[main_legend_index]]
    # Create a text grob for the main legend title.
    # Set x to the center, justify the text to "center", and force the width to fill the space.
    main_title_grob <- grid::textGrob("Scaled Normalized Counts",
                                      gp = grid::gpar(fontsize = 10, fontface = "bold"),
                                      x = unit(0, "npc"), just = "left",
                                      name = "main_legend_title")
    
    # If the main legend is not already a gtable, wrap it in one.
    if (!inherits(main_legend, "gtable")) {
      tmp_main <- main_legend
      main_legend <- gtable::gtable(widths = unit(5, "npc"), heights = unit(1, "npc"))
      main_legend <- gtable::gtable_add_grob(main_legend, tmp_main, t = 1, l = 1)
    }
    
    # Add a new row at the top of the main legend for the title.
    # Here we add extra space (unit(10, "mm")) to help display long titles.
    main_legend <- gtable::gtable_add_rows(main_legend,
                                           heights = grid::grobHeight(main_title_grob) + unit(5, "mm"),
                                           pos = 0)
    main_legend <- gtable::gtable_add_grob(main_legend, main_title_grob,
                                           t = 1, l = 1, r = ncol(main_legend))
    
    
    # Add an extra column at the left to shift the legend to the right by 2 mm.
    main_legend <- gtable::gtable_add_cols(main_legend, widths = unit(45, "mm"), pos = 0)

    # Replace the original main legend with the modified version
    g$grobs[[main_legend_index]] <- main_legend
  }
  
  # Locate the annotation legend (its layout name is "annotation_legend")
  anno_legend_index <- which(g$layout$name == "annotation_legend")
  if (length(anno_legend_index) > 0) {
    # Extract the annotation legend grob
    anno_legend <- g$grobs[[anno_legend_index]]
    
    # Remove the annotation legend from its default position
    g$grobs[[anno_legend_index]] <- grid::nullGrob()
    
    # Add a new row at the bottom of the overall gtable to accommodate the annotation legend.
    # (We add a little extra spacing with unit(2, "mm").)
    g <- gtable::gtable_add_rows(g,
                                 heights = grid::grobHeight(anno_legend) + unit(20, "mm"))
    
    # Place the annotation legend into the new bottom row spanning the full width.
    g <- gtable::gtable_add_grob(g, anno_legend,
                                 t = nrow(g), l = 1, r = ncol(g))
  }
  
  
  
  
  
  
  return(g)
  # Draw the modified heatmap
  #grid::grid.newpage()
  #grid::grid.draw(g)
  
  invisible(ph)
}

```


```{r}
# color_setting

if (!file.exists(file.path(COUNT_path,"Biobank-samples-map-clinical-102523.xlsx"))){
  synapser::synLogin(email="",token="")


# DOWNLOAD CLINICAL METADATA (uncomment next line to download file)
 synapser::synGet("syn52798607", downloadLocation = COUNT_path,)
}



# READ IN OFFLINE SAMPLESHEET
dir(COUNT_path)
jhu_clinicaldata_1 <- readxl::read_xlsx(file.path(COUNT_path,"Biobank-samples-map-clinical-102523.xlsx")) %>% 
  dplyr::filter(assay == "Whole Exome Sequencing") %>%  #select WES samples
  dplyr::filter(batch != "WU_batch3") %>%  # select only batch1, 2, 3 samples
  dplyr::filter(tissue == "primary tumor") %>%  # select only tumor samples
  dplyr::mutate(Aliquot_ID_Synapse_NEW = tidyr::replace_na(Aliquot_ID_Synapse_NEW, "A")) %>% 
  dplyr::mutate(Tumor_Sample_Barcode = stringr::str_c(SpecimenID_Synapse_NEW, "-", Aliquot_ID_Synapse_NEW)) %>% 
  dplyr::mutate(
    # Create categories
    age_group = dplyr::case_when(
      age <= 10            ~ "0-10",
      age > 10 & age <= 20 ~ "11-20",
      age > 20 & age <= 30 ~ "21-30",
      age > 30 & age <= 40 ~ "31-40",
      age > 40 & age <= 50 ~ "41-50",
      age > 50             ~ "> 50"
    ),
    # Convert to factor
    age_group = factor(
      age_group,
      level = c("0-10", "11-20","21-30", "31-40", "41-50", "> 50")
    )
  ) %>% 
  dplyr::select(tumorType, sex, age_group, location, tissue, individualID, Tumor_Sample_Barcode, NF1_loss_clinical_sequencing, NF1_loss_clinical_diagnosis) %>% 
  dplyr::filter(!(Tumor_Sample_Barcode %in% c("JH-2-015-6AD96-A", "JH-2-106-B73E8-A", "JH-2-106-1447G-A", "JH-2-084-F9A84-A", "JH-2-084-AE169-A"))) %>% #remove 2-015TD, JH-2-106, JH-2-084 
  purrr::set_names(nm = c(
    "tumorType", 
    "sex", 
    "age_group", 
    "location", 
    "tissue", 
    "individualID", 
    "Tumor_Sample_Barcode", 
    "NF1_clinical_sequencing", 
    "NF1_clinical_diagnosis"
  ))

load(file.path(Process_data,"MPNST_VS_PN_Deseq2_outut.rdata"))

```



```{r colors, echo=TRUE, eval=TRUE, results='hide', message=FALSE, warning=FALSE}

## Assign colors to the metadata for plotting 
# Define colors for tumor types
#tumortype_colors <- c("#0D3B66", "#2d728f","#ab3428","#61C9A8","#7E52A0", "#F4D35E")
tumortype_colors <- c("#0D3B66", "#e8be00","#ab3428","#61C9A8","#7E52A0", "#CABACB")

names(tumortype_colors) <- jhu_clinicaldata_1$tumorType %>%unique()

tumortype_colors=tumortype_colors[c(1,2,3,4,6)]

# PCA plot for all tumors
set.seed(123456)
Tumor_color=tumortype_colors%>%na.omit()

mat_colors <- list(Group = Tumor_color[1:2])
names(mat_colors$Group) <- c("PN","MPNST")

Manu_color=unlist(mat_colors)
names(Manu_color)=gsub("Group.", "",names(Manu_color))

Tumor_color2=tumortype_colors%>%na.omit()
names(Tumor_color2)=tolower(names(Tumor_color2))

mat_colors <- list(Group = Tumor_color[1:2])
names(mat_colors$Group) <- c("PN","MPNST")

Manu_color=unlist(mat_colors)
names(Manu_color)=gsub("Group.", "",names(Manu_color))


```



```{r}
# read sample annotation file
Disease=data.frame(tumorType=c("Plexiform Neurofibroma","Cutaneous Neurofibroma", "Diffuse Infiltrating Neurofibroma" ,"Nodular Neurofibroma" ,"Malignant Peripheral Nerve Sheath Tumor"))%>%mutate(Disease=c("PN","cNF","DifNF","NodNF","MPNST"  ))

RNAseq_anno=read.csv(file.path(COUNT_path,"RNAseq_samplemap.csv"))%>%mutate(batch=ifelse(specimenID=="JH-2-023-91HBD","WU_batch2",batch))%>%mutate(Sample_name=Syn_speicimenID)%>%left_join(.,Disease)%>%mutate("Tumor_type"=tolower(tumorType))
RNAseq_anno$tumorType=factor(RNAseq_anno$tumorType,levels = names(Tumor_color))
RNAseq_anno$"Tumor_type"=factor(RNAseq_anno$"Tumor_type",levels = tolower(names(Tumor_color)))


# load data from Deseq2
MPNST_PN_RNAseq_anno=RNAseq_anno%>%dplyr::filter(Disease%in%c("MPNST","PN"))
Expression_vst_anno_MPNST_PN=Expression_vst_anno%>%dplyr::filter(Sample_name%in%MPNST_PN_RNAseq_anno$Sample_name)
Expression_vst_MPNST_PN=Expression_vst[,rownames(Expression_vst_anno_MPNST_PN)]
MPNST_vs_PN_res_2=MPNST_vs_PN_res%>%mutate(gene_symbol=rownames(.))
MPNST_vs_PN_res_q005=MPNST_vs_PN_res %>%dplyr::filter(padj<0.05)

q005_gene=rownames(MPNST_vs_PN_res_q005)
Expression_vst_anno_MPNST_PN_q005=Expression_vst_anno_MPNST_PN[,c("Sample_name",q005_gene)]
Expression_vst_MPNST_PN_q005=Expression_vst_MPNST_PN[q005_gene,]

MPNST_PN_sample=MPNST_PN_RNAseq_anno [,c("Sample_name","individualID","tumorType","Disease")]



#DEGs list
sig_gene=MPNST_vs_PN_res %>%dplyr::filter(padj<0.05)%>%rownames(.)
Expression_vst_MPNST_PN_sig=Expression_vst_MPNST_PN[sig_gene,]
MPNST_vs_PN_DEGs=MPNST_vs_PN_res_2%>%dplyr::filter(padj<0.05&(abs(log2FoldChange)>0.27) )
```


\newpage


# Figure 4

```{r,include=F}
# Fig4A
P1=PCA_Plot_3  (t(Expression_vst),RNAseq_anno,"Sample_name","Tumor_type",Tumor_color2)
# UMAP plot for all tumors
#P2=UMAP_plot(t(Expression_vst),RNAseq_anno)+scale_color_manual(values = Tumor_color)

#cowplot::plot_grid(P1,"",rel_widths = c(1,1.5),labels = c("A","B"))
```

# MPNST assoicated genes
**S100B, CNP,PMP22,NGFR,SOX10, ,Expression of Schwann cell differentiation markers Downregulaed in MPNST** (PMID: 16510576)
**TWIST1,SOX9 ,stem cell marker, upregulated in MPNST"**  (PMID: 16510576)
**E-cadherin (CDH1) was downregulated in both neurofibroma and MPNSTs**
**GAS1 (Growth Arrest Specific 1) is a Protein Coding gene. IFG2, PTK7, FGFR1, TWIST1, GAS1, and EGFR.upregulated in MPNST** (PMID: 32642732)
** UBR5 PMID: 39857943**

```{r,include=F,fig.width=18,fig.height=12}
# Fig4B (11 mpnst genes DEGS padj0.05 FC1.2)
# Define MPNST assoicated genes
MPNST_associated_genes= c( "UBR5","TP53", "CNP","PMP22","NGFR","SOX9","SOX10","S100B","S100A4", "S100A1", "S100A2", "S100A6","CCNB2","TWIST1","EGFR","FGFR1","ERBB3","ERBB2","COL6A3","CDH1","CDKN2A","PDGFA","PDGFB","TGFBR2","GAS1","TGFB1","TGFB2","IGF2","PTK7")

length(MPNST_associated_genes)


Pathway_list=list("MPNST_associated_genes"=MPNST_associated_genes)
Stat_method="wilcox.test"
gene_list2=data.frame(MPNST_associated_genes=MPNST_associated_genes)
colnames(Expression_vst_MPNST_PN_q005)==MPNST_PN_sample$Sample_name


MPNST_associated_genes_DEGs=MPNST_vs_PN_DEGs%>%dplyr::filter(gene_symbol%in%MPNST_associated_genes)%>%arrange(log2FoldChange)
# P2=Volcano_plot_rnaseq_add_label(MPNST_vs_PN_res_2,Pathway_list,"MPNST associated genes")
#P2=Singscore_data_plot_2(gene_list2,Expression_vst_MPNST_PN_q005,Expression_vst_anno_MPNST_PN_q005,GTAC_Sample_anno,Stat_method,mat_colors)

mat_colors <- list(Tumor_type = Tumor_color2[1:2])
#names(mat_colors$Group) <- c("PN","MPNST")

Anno_table=RNAseq_anno[,c("Sample_name","Tumor_type")]  %>%as.data.frame()%>%dplyr::filter(Sample_name%in%colnames(Expression_vst_MPNST_PN_q005))%>%tibble::column_to_rownames(.,"Sample_name")

```

```{r,include=T}
knitr::kable(MPNST_associated_genes_DEGs,caption = "MPNST associated genes (DEGS)")
```





```{r,include=T}
MPNST_PN_sample_2=RNAseq_anno%>%dplyr::filter(Disease%in%c("MPNST","PN"))%>%.[,c("Sample_name","individualID","batch","Disease")]%>%arrange(Disease)
knitr::kable(MPNST_PN_sample_2,caption="PN and MPNST")%>%kableExtra::kable_styling( latex_options="scale_down" )
```

\newpage

```{r}
P2=pheatmap_pathway_annotation_bottom(expression_data=t(Expression_vst_MPNST_PN_q005),Gene_select=MPNST_associated_genes,pathway_name=names(Pathway_list), Annotation=Anno_table,anno_color=mat_colors)

```


```{r,include=T,fig.width=20,fig.height=16}
cowplot:: plot_grid(P1,grid::nullGrob(),nrow = 1,labels = c("A","B"),scale = c(0.4,1),label_size = 20,vjust=1)
#cowplot:: plot_grid(P1,grid::nullGrob(), P2,nrow = 2,labels = c("A","","B"),rel_heights = c(0.6,1,1),label_size = 20,vjust=1)

```

```{r,include=T,fig.width=30,fig.height=24}
cowplot:: plot_grid(P2,grid::nullGrob(),nrow = 1,labels = c("B","V"),scale = c(1.2,1),label_size = 20,vjust=1)
```

```{r,eval=F,fig.cap="MPNST markers expression box plot",include=T,fig.height=20,fig.width=14,out.height="95%"}
# Compare_list=list(c("MPNST","PN"))
# setdiff(MPNST_associated_genes,colnames(Expression_vst_anno_MPNST_PN))
# 
# res_2=MPNST_vs_PN_res%>%mutate(group1="MPNST",group2="PN")
```


# GSEA analysis (The combination of Hallmark and C2 genesets  )
```{r,include=F}
GMT_files=list.files(Reference_path,pattern="2023.2",full.names = T)%>%.[grepl("gmt.txt",.)]%>%.[grepl("c2.|h",.)]
pathway_all= c(fgsea::gmtPathways(GMT_files[1]),fgsea::gmtPathways(GMT_files[2]))
res_q005_rank=MPNST_vs_PN_res_q005%>%arrange(log2FoldChange) %>%.[c("log2FoldChange")]%>%as.data.frame() %>%tibble::rownames_to_column()
ranks <- tibble::deframe(res_q005_rank)

set.seed(1234) # keep set seed in the front of fgsea loop otherwise it will change the result
fgseaRes_all <- fgsea::fgseaMultilevel(pathway_all, ranks,
                           minSize=5,
                           maxSize=500
                           )
fgseaRes_all%>%dplyr::filter(grepl("WP",pathway))

Fgsea_all_sig_2=fgseaRes_all%>%dplyr::filter(padj<0.05)%>%arrange(padj)%>%dplyr::rename(gene_counts=size)
#Hallmark_2=Fgsea_all_sig_2%>%dplyr::filter(grepl("HALLMARK",pathway))
write.csv(Fgsea_all_sig_2[,c("pathway","padj","NES","gene_counts")],file.path(Process_data,"sig_combined_pathways_in_MPNST_vs_PN.csv"),row.names = F)

```

\newpage

# Top 30 pathways (filter pathways by NES>0, sort by adj, choose top 30 )

```{r,include=F}
Fgsea_all_sig_2$pathway=gsub("_"," ",Fgsea_all_sig_2$pathway)
Fgsea_all_sig_2_up=Fgsea_all_sig_2%>%dplyr::filter(NES>0)
Fgsea_all_sig_2_DN=Fgsea_all_sig_2%>%dplyr::filter(NES<0)
write.csv(Fgsea_all_sig_2_up[,c("pathway","padj","NES","gene_counts")],file.path(Process_data,"sig_combined_up_pathways_in_MPNST_vs_PN.csv"),row.names = F)
write.csv(Fgsea_all_sig_2_DN[,c("pathway","padj","NES","gene_counts")],file.path(Process_data,"sig_combined_DN_pathways_in_MPNST_vs_PN.csv"),row.names = F)


Fgsea_all_sig_2_up_top30=Fgsea_all_sig_2_up[,c("pathway","padj","NES","gene_counts")][1:30,]%>%arrange(desc(NES))

write.csv(Fgsea_all_sig_2_up_top30[,c("pathway","padj","NES","gene_counts")],file.path(Process_data,"sig_30_activated_pathways_in_MPNST_vs_PN.csv"),row.names = F)

```

```{r,include=T}
library(kableExtra)
#knitr::kable(Fgsea_all_sig_2_up_top30,caption = "Top 30 activated pathways (NES>0,top 30 by padj)")%>%kableExtra::kable_styling(latex_options = c("scale_down"),full_width=F,font_size=5)%>%column_spec(1, width = "5cm",latex_column_spec = "p{5cm}") 

knitr::kable(
  Fgsea_all_sig_2_up_top30,
  caption = "Top 30 activated pathways (NES > 0, top 30 by padj)"
) |>
  kableExtra::kable_styling(
    latex_options = "scale_down",
    full_width = FALSE,
    font_size = 5  # or any size in points
  )

```

\newpage

```{r,include=F}
P7=Enrich_dotplot(Fgsea_all_sig_2_up[1:30,],31,size_max=25,text_y_size = 60) 

```


```{r,include=T,fig.width=60,fig.height=70,out.width="70%"}
P7
```
```{r}
Chosen_pathways=c("MEISSNER_BRAIN_HCP_WITH_H3K27ME3","HALLMARK_EPITHELIAL_MESENCHYMAL_TRANSITION")
Other_list=list()

for (i in 1:length(Chosen_pathways)){
  Name=Chosen_pathways[i]
  
  #Pathway_list=pathway[[Name]]%>%as.list()
  #names(Pathway_list)=Name
  Other_list[[i]]=Volcano_plot_rnaseq_add_label_update_2(MPNST_vs_PN_res_2,pathway_all,Name,top_gene_number = 10,New_title=gsub("_"," ",Name))
  
}

#install.packages("openxlsx")
 


for (i in 1:length(Chosen_pathways)){
  Name=Chosen_pathways[i]%>%gsub(" ","_",.)
  Pathway_gene_list=pathway_all[[Name]]
  DEGs_pathway=MPNST_vs_PN_DEGs%>%dplyr::filter(gene_symbol%in%Pathway_gene_list)%>%arrange(desc(log2FoldChange))%>%.[,c("log2FoldChange","pvalue","padj")]
  write.csv(DEGs_pathway,file=file.path(Process_data,paste0(Name,"_DEGs_padj005_FC0.27.csv")))
}
  

# heatmap_list=list()
# 
# for (i in 1:length(Chosen_pathways)){
#      Name=Chosen_pathways[i]
#      Pathway_list=pathway_all[[Name]]%>%as.data.frame()%>%setNames(Name)
#      Stat_method="wilcox.test"
#      heatmap_list[[i]]=Singscore_data_plot_2(Pathway_list,Expression_vst_MPNST_PN_q005,Expression_vst_anno_MPNST_PN_q005,GTAC_Sample_anno,Stat_method,mat_colors)[[1]]
# }
# 
# heatmap_list[[2]]
```



```{r,fig.width=22,fig.height=12,include=T,out.width="60%"}
cowplot::plot_grid(plotlist = Other_list,nrow=1,labels = c("D"))
```


# Supplementary files

```{r}
Chosen_pathways=Fgsea_all_sig_2_up$pathway[1:5]

Other_list=list()

for (i in 1:length(Chosen_pathways)){
  Name=Chosen_pathways[i]%>%gsub(" ","_",.)
  
  #Pathway_list=pathway_all[[Name]]%>%as.list()
  #names(Pathway_list)=Name
  Other_list[[i]]=Volcano_plot_rnaseq_add_label_update_2(MPNST_vs_PN_res_2,pathway_all,Name,top_gene_number = 10,New_title=gsub("_"," ",Name))
  
}
```


```{r}
for (i in 1:length(Chosen_pathways)){
  Name=Chosen_pathways[i]%>%gsub(" ","_",.)
  Pathway_gene_list=pathway_all[[Name]]
  DEGs_pathway=MPNST_vs_PN_DEGs%>%dplyr::filter(gene_symbol%in%Pathway_gene_list)%>%arrange(desc(log2FoldChange))%>%.[,c("log2FoldChange","pvalue","padj")]
  write.csv(DEGs_pathway,file=file.path(Process_data,paste0(Name,"_DEGs_padj005_FC0.27.csv")))
}
```


```{r}
library(dplyr)
library(openxlsx)

# Create a new workbook
wb <- createWorkbook()

for (i in seq_along(Chosen_pathways)) {
  Name <- gsub(" ", "_", Chosen_pathways[i])
  Pathway_gene_list <- pathway_all[[Name]]
  
  DEGs_pathway <- MPNST_vs_PN_DEGs %>%
    filter(gene_symbol %in% Pathway_gene_list) %>%
    arrange(desc(log2FoldChange)) %>%
    select(gene_symbol,log2FoldChange, pvalue, padj)
  
  # Add a worksheet and write data
  addWorksheet(wb, sheetName = i)
  title_text <- paste0("Significant differentially expressed genes in ", Name,
                       " between MPNST and PNF (padj <0.05 & |log2FoldChange| > 0.27)")
  writeData(wb, sheet = i, x = data.frame(title_text), startRow = 1, colNames = FALSE)
  writeData(wb, sheet = i, x = DEGs_pathway, startRow = 3, colNames = TRUE)
 
}

# Save workbook
saveWorkbook(wb, file = file.path(Process_data, "Top_5_Pathways_DEGs.xlsx"), overwrite = TRUE)

```


```{r,fig.width=28,fig.height=22,include=T}

cowplot::plot_grid(plotlist = Other_list,ncol=3)
```


```{r,include=F,fig.height=6,fig.width=8}
top_heatmap_list=list()

for (i in Fgsea_all_sig_2_up$pathway[1:5]){
    i=gsub(" ","_",i)
    Pathway_list=pathway_all[[i]]
    #%>%as.data.frame()%>%setNames(i)
    Stat_method="wilcox.test"
# gene_list2=GSEA_pathway_list[[Pathway]]%>%as.data.frame() %>%setNames(Pathway)

    top_heatmap_list[[i]]=pheatmap_pathway_annotation_bottom(expression_data=t(Expression_vst_MPNST_PN_q005),Gene_select=Pathway_list,pathway_name=gsub("_"," ",i), Annotation=Anno_table,anno_color=mat_colors)
}
```


```{r,fig.width=14,fig.height=16,include=T}
cowplot::plot_grid(top_heatmap_list[[1]])

```



```{r,fig.width=32,fig.height=40,include=T}
cowplot::plot_grid( top_heatmap_list[[2]])

```

```{r,fig.width=28,fig.height=40,include=T}
cowplot::plot_grid(top_heatmap_list[[3]])

```


```{r,fig.width=36,fig.height=54,include=T}
cowplot::plot_grid( top_heatmap_list[[4]])

```

```{r,fig.width=14,fig.height=16,include=T}
cowplot::plot_grid(top_heatmap_list[[5]])

```





```{r}
knitr::knit_exit()
```


\newpage

## Reactome pathway:Fig s2
```{r,include=F}
Reactome_pathway= Fgsea_all_sig%>%dplyr::filter(grepl("REACTOME",pathway))%>%.[,c(1,3,6,7)]%>%arrange(desc(NES))

```


```{r,include=F}
P6=Enrich_dotplot(Reactome_pathway,40,size_max=15)
```


```{r,include=T,fig.width=26,fig.height=16,out.width="80%"}
cowplot::plot_grid(P6)
```


```{r,include=F,fig.width=20,fig.height=15}
Reactome_list=list()
# Reactome_pathway_2=Reactome_pathway%>%dplyr::filter(pathway%in%c("REACTOME_DEGRADATION_OF_THE_EXTRACELLULAR_MATRIX","REACTOME_EXTRACELLULAR_MATRIX_ORGANIZATION","REACTOME_CELL_CYCLE_MITOTIC","REACTOME_EGR2_AND_SOX10_MEDIATED_INITIATION_OF_SCHWANN_CELL_MYELINATION"))
Reactome_pathway=Reactome_pathway%>%arrange(desc(NES))
Pathway_select=c("REACTOME_COLLAGEN_DEGRADATION","REACTOME_EXTRACELLULAR_MATRIX_ORGANIZATION","REACTOME_DEGRADATION_OF_THE_EXTRACELLULAR_MATRIX")
for (i in Pathway_select){
  Name=i
  Reactome_list[[i]]=Volcano_plot_rnaseq_add_label_update_2(MPNST_vs_PN_res_2,pathway_all,Name,top_gene_number = 10,New_title=gsub("_"," ",Name))
  
}
```


```{r,include=T,fig.width=36,fig.height=15}
cowplot::plot_grid( plotlist=Reactome_list,ncol=3 ) 
```

\newpage

```{r,include=F,fig.height=6,fig.width=8}
heatmap_list=list()
for (i in 1:length(Pathway_select)){
    Name=Pathway_select[i]
    Pathway_list=pathway_all[[Name]]
    #%>%as.data.frame()%>%setNames(i)
    Stat_method="wilcox.test"
# gene_list2=GSEA_pathway_list[[Pathway]]%>%as.data.frame() %>%setNames(Pathway)

    heatmap_list[[i]]=pheatmap_pathway_annotation_bottom(expression_data=t(Expression_vst_MPNST_PN_q005),Gene_select=Pathway_list,pathway_name=Name, Annotation=Anno_table,anno_color=mat_colors)
}
```


```{r,fig.width=14,fig.height=16,include=T}
cowplot::plot_grid(heatmap_list[[1]])

```

\newpage
```{r,fig.width=14,fig.height=16,include=T}
cowplot::plot_grid(heatmap_list[[2]])

```

\newpage
```{r,fig.width=14,fig.height=16,include=T}
cowplot::plot_grid(heatmap_list[[3]])

```



\newpage

# Fig S3

```{r,include=F,fig.width=20,fig.height=15}
Other_list=list()
# Reactome_pathway_2=Reactome_pathway%>%dplyr::filter(pathway%in%c("REACTOME_DEGRADATION_OF_THE_EXTRACELLULAR_MATRIX","REACTOME_EXTRACELLULAR_MATRIX_ORGANIZATION","REACTOME_CELL_CYCLE_MITOTIC","REACTOME_EGR2_AND_SOX10_MEDIATED_INITIATION_OF_SCHWANN_CELL_MYELINATION"))
Three_pathways=c("MEISSNER_BRAIN_HCP_WITH_H3K27ME3","WINNEPENNINCKX_MELANOMA_METASTASIS_UP","HALLMARK_MTORC1_SIGNALING")

names(pathway)%>%.[Three_pathways[3]]

setdiff(names(pathway)%>%.[grepl(Three_pathways[1],.)],Three_pathways)
pathway[["MEISSNER_BRAIN_HCP_WITH_H3K27ME3"]]
pathway[[Three_pathways[1]]]

for (i in 1:length(Three_pathways)){
  Name=Three_pathways[i]
  Other_list[[i]]=Volcano_plot_rnaseq_add_label_update_2(MPNST_vs_PN_res_2,pathway_all,Name,top_gene_number = 10,New_title=gsub("_"," ",Name))
  
  
}
```

```{r,include=T,fig.width=36,fig.height=15}
cowplot::plot_grid(plotlist = Other_list,nrow=1,labels = c("A","B","C"),label_size = 30)
```

```{r}
outlist=list()
for (i in 1:length(Three_pathways)){
  Name=Three_pathways[i]
  Gene_list=pathway_all[Name]
  outlist[[i]]=Singscore_boxplot(Gene_list,Expression_vst_MPNST_PN_q005,RNAseq_anno,"Tumor_type",Stat_method,Tumor_color2[1:2])
}

```

```{r,fig.height=14,fig.width=20,include=T}
cowplot::plot_grid(plotlist = outlist,nrow=1,labels = c("D","E","F"),label_size = 30,vjust=-0.5)
```



```{r,include=F,fig.height=6,fig.width=8}
heatmap_list=list()
for (i in 1:length(Three_pathways)){
    Name=Three_pathways[i]
    Pathway_list=pathway_all[[Name]]
    #%>%as.data.frame()%>%setNames(i)
    Stat_method="wilcox.test"
# gene_list2=GSEA_pathway_list[[Pathway]]%>%as.data.frame() %>%setNames(Pathway)

    heatmap_list[[i]]=pheatmap_pathway_annotation_bottom(expression_data=t(Expression_vst_MPNST_PN_q005),Gene_select=Pathway_list,pathway_name=Name, Annotation=Anno_table,anno_color=mat_colors)
}
```


```{r,fig.width=14,fig.height=16,include=T}
cowplot::plot_grid(heatmap_list[[1]])

```



```{r,fig.width=14,fig.height=16,include=T}
cowplot::plot_grid(heatmap_list[[2]])

```

```{r,fig.width=14,fig.height=16,include=T}
cowplot::plot_grid(heatmap_list[[3]])

```





```{r}
knitr::knit_exit()
```






```{r,include=F,fig.width=20,fig.height=15}
Other_heatmap_list=list()

for (i in 1:length(Three_pathways)){
    Name=Three_pathways[i]
    Pathway_list=pathway[[Name]]%>%as.data.frame()%>%setNames(Name)
    Stat_method="wilcox.test"
    Other_heatmap_list[[i]]=Singscore_data_plot_2(Pathway_list,Expression_vst_MPNST_PN_q005,Expression_vst_anno_MPNST_PN_q005,GTAC_Sample_anno,Stat_method,mat_colors)[[1]]$gtable
}
```


```{r}
knitr::knit_exit()
```



```{r,eval=F}
Permutation_test=function(data,genelist) {
 
  ranked <- singscore::rankGenes(data)
  scoredf <- singscore::simpleScore(ranked, upSet = genelist)
 # find out what backends can be registered on your machine
  
  BiocParallel::registered()
  permuteResult = singscore::generateNull(upSet = genelist, downSet = NULL,rankData= ranked,B =100,   seed = 1234, useBPPARAM = NULL)
   # call the permutation function to generate the empirical scores
 # for B times.
 pvals <- singscore::getPvals(permuteResult,scoredf)
  return(list(permuteResult,scoredf,pvals))
}


Permu_out=Permutation_test(Expression_vst_MPNST_PN, Gene_list3)
```




```{r}
Permutation_test=function(data,genelist) {
 
  ranked <- singscore::rankGenes(data)
  scoredf <- singscore::simpleScore(ranked, upSet = genelist)
 # find out what backends can be registered on your machine
  
  BiocParallel::registered()
  permuteResult = singscore::generateNull(upSet = genelist, downSet = NULL,rankData= ranked,B =3000,   seed = 1234, useBPPARAM = NULL)
   # call the permutation function to generate the empirical scores
 # for B times.
 pvals <- singscore::getPvals(permuteResult,scoredf)
  return(list(permuteResult,scoredf,pvals))
}



```



```{r,include=F,fig.width=6,fig.height=18}
MPNST_vs_PN_res_q005_up=MPNST_vs_PN_res_q005%>%dplyr::filter(log2FoldChange>0)
#MPNST_vs_q005_up=Expression_vst_MPNST_PN_q005[rownames(MPNST_vs_PN_res_q005_up),MPNST_samples]

Permu_out=Permutation_test(Expression_vst_MPNST_PN, Gene_list3)
MPNST_samples=RNAseq_anno%>%dplyr::filter(Disease=="MPNST")%>%.$Syn_speicimenID
PN_samples=RNAseq_anno%>%dplyr::filter(Disease=="PN")%>%.$Syn_speicimenID
Permu_out[[3]]
singscore::plotNull(Permu_out[[1]], Permu_out[[2]], Permu_out[[3]], sampleNames = MPNST_samples)
```

```{r}
data_color=data.frame(Sample_name=rownames(Permu_out[[2]]))%>%left_join(RNAseq_anno)
Data_input=Permu_out[[2]]%>%mutate(Disease=data_color$Disease)
singscore::plotDispersion(Data_input)+ggtitle("test")+
  geom_point(aes(color =Disease))
```


```{r}
ranked <- singscore::rankGenes(as.data.frame(Expression_vst_MPNST_PN))%>%as.data.frame()
```


```{r}
PN_plot_list=list()
for (i in PN_samples) {
  input=ranked[i]
  PN_plot_list[[i]]=singscore::plotRankDensity(input,upSet=Gene_list2)+ggtitle(i)
}

 
MPNST_plot_list=list()
for (i in MPNST_samples) {
  input=ranked[i]
  MPNST_plot_list[[i]]=singscore::plotRankDensity(input,upSet=Gene_list2)+ggtitle(i)
}
```

```{r,include=F,fig.width=8,fig.height=20}
cowplot::plot_grid(plotlist =PN_plot_list,ncol=1 )
```


```{r,include=F,fig.width=8,fig.height=20}
cowplot::plot_grid(plotlist =MPNST_plot_list,ncol=1 )
```


```{r,include=F,fig.width=6,fig.height=18}
singscore::plotNull(Permu_out[[1]], Permu_out[[2]], Permu_out[[3]], sampleNames = PN_samples)


```
 
 
```{r,include=F,fig.width=6,fig.height=24}

PN_samples=RNAseq_anno%>%dplyr::filter(Disease=="PN")%>%.$Syn_speicimenID

singscore::plotNull(Permu_out[[1]], Permu_out[[2]], Permu_out[[3]], sampleNames = PN_samples)
```


```{r,include=F,fig.width=20,fig.height=15}
names(pathway)%>%.[Three_pathways[3]]

setdiff(names(pathway)%>%.[grepl(Three_pathways[1],.)],Three_pathways)
pathway[["MEISSNER_BRAIN_HCP_WITH_H3K27ME3"]]
pathway[[Three_pathways[1]]]

for (i in 1:length(Three_pathways)){
  Name=Three_pathways[i]
  Pathway_list=pathway[[Name]]%>%as.list()
  names(Pathway_list)=Name
  Other_list[[i]]=Volcano_plot_rnaseq_add_label_update(MPNST_vs_PN_res_2,pathway,Name)
  
}


Other_heatmap_list=list()

for (i in 1:length(Three_pathways)){
    Name=Three_pathways[i]
    Pathway_list=pathway[[Name]]%>%as.data.frame()%>%setNames(Name)
    Stat_method="wilcox.test"
    Other_heatmap_list[[i]]=Singscore_data_plot_2(Pathway_list,Expression_vst_MPNST_PN_q005,Expression_vst_anno_MPNST_PN_q005,GTAC_Sample_anno,Stat_method,mat_colors)[[1]]$gtable
}
```
 



## Reactome pathway
```{r,include=T}
Reactome_pathway= Fgsea_all_sig%>%dplyr::filter(grepl("REACTOME",pathway))%>%.[,c(1,3,6,7)]%>%arrange(desc(NES))

```


```{r,fig.width=18,fig.height=12,include=F}

P6=Enrich_dotplot(Reactome_pathway,45,split=F)
```


```{r,include=T,fig.width=18,fig.height=15}
Hallmark_list=list()

for (i in Hallmark_pathway$pathway[1:3]){
  Pathway_list=pathway[[i]]%>%as.list()
  names(Pathway_list)=i
  Hallmark_list[[i]]=Volcano_plot_rnaseq_add_label_update(MPNST_vs_PN_res_2,pathway,i)
  
}

P7=cowplot::plot_grid( plotlist =Hallmark_list,ncol=3   ) 
```


```{r,include=T,fig.width=20,fig.height=15}

Reactome_list=list()
# Reactome_pathway_2=Reactome_pathway%>%dplyr::filter(pathway%in%c("REACTOME_DEGRADATION_OF_THE_EXTRACELLULAR_MATRIX","REACTOME_EXTRACELLULAR_MATRIX_ORGANIZATION","REACTOME_CELL_CYCLE_MITOTIC","REACTOME_EGR2_AND_SOX10_MEDIATED_INITIATION_OF_SCHWANN_CELL_MYELINATION"))


for (i in Reactome_pathway$pathway[1:3]){
  Pathway_list=pathway[[i]]%>%as.list()
  names(Pathway_list)=i
  Reactome_list[[i]]=Volcano_plot_rnaseq_add_label_update(MPNST_vs_PN_res_2,pathway,i)
  
}

P8=cowplot::plot_grid( plotlist =Reactome_list,ncol=3   ) 
```


\newpage

# Figure 4

```{r,fig.width=28,fig.height=16,include=T}
cowplot:: plot_grid(P5,P7,P6,P8,nrow = 2,labels = c("A","B","C","D"),rel_widths = c(0.6,1,0.7,1))

```

\newpage

# Supplymentary tables


```{r,include=T}
knitr::kable(Hallmark_pathway,caption = "Significant Hallmark pathways in MPNST vs PN (Adjust p value <0.05)")%>%kableExtra::kable_styling(latex_options = c("hold_position"))


knitr::kable(Reactome_pathway,caption = "Significant Reactome pathways in MPNST vs PN (Adjust p value <0.05)")%>%kableExtra::kable_styling(latex_options = c("hold_position","scale_down"))
```

\newpage

# Supplymentary fig1






```{r,include=F,fig.height=6,fig.width=8}
Reactome_heatmap_list=list()

for (i in Reactome_pathway$pathway[1:3]){
    Pathway_list=pathway[[i]]%>%as.list()
    names(Pathway_list)=i
    Pathway_list=pathway[[i]]%>%as.data.frame()%>%setNames(i)
    Stat_method="wilcox.test"
# gene_list2=GSEA_pathway_list[[Pathway]]%>%as.data.frame() %>%setNames(Pathway)

 
    Reactome_heatmap_list[[i]]=Singscore_data_plot_2(Pathway_list,Expression_vst_MPNST_PN_q005,Expression_vst_anno_MPNST_PN_q005,GTAC_Sample_anno,Stat_method,mat_colors)[[1]]$gtable
}
```



```{r,include=T,fig.height=34,fig.width=22}

cowplot::plot_grid(plotlist = c(Hallmark_heatmap_list,Reactome_heatmap_list),ncol=3,scale= 1)

```


\newpage 

# Supplementary fig2


```{r,include=F,fig.width=20,fig.height=15}
Other_list=list()
# Reactome_pathway_2=Reactome_pathway%>%dplyr::filter(pathway%in%c("REACTOME_DEGRADATION_OF_THE_EXTRACELLULAR_MATRIX","REACTOME_EXTRACELLULAR_MATRIX_ORGANIZATION","REACTOME_CELL_CYCLE_MITOTIC","REACTOME_EGR2_AND_SOX10_MEDIATED_INITIATION_OF_SCHWANN_CELL_MYELINATION"))
Three_pathways=c("MEISSNER_BRAIN_HCP_WITH_H3K27ME3","WINNEPENNINCKX_MELANOMA_METASTASIS_UP","HALLMARK_MTORC1_SIGNALING")

names(pathway)%>%.[Three_pathways[3]]

setdiff(names(pathway)%>%.[grepl(Three_pathways[1],.)],Three_pathways)
pathway[["MEISSNER_BRAIN_HCP_WITH_H3K27ME3"]]
pathway[[Three_pathways[1]]]

for (i in 1:length(Three_pathways)){
  Name=Three_pathways[i]
  Pathway_list=pathway[[Name]]%>%as.list()
  names(Pathway_list)=Name
  Other_list[[i]]=Volcano_plot_rnaseq_add_label_update(MPNST_vs_PN_res_2,pathway,Name)
  
}


Other_heatmap_list=list()

for (i in 1:length(Three_pathways)){
    Name=Three_pathways[i]
    Pathway_list=pathway[[Name]]%>%as.data.frame()%>%setNames(Name)
    Stat_method="wilcox.test"
    Other_heatmap_list[[i]]=Singscore_data_plot_2(Pathway_list,Expression_vst_MPNST_PN_q005,Expression_vst_anno_MPNST_PN_q005,GTAC_Sample_anno,Stat_method,mat_colors)[[1]]$gtable
}
```


```{r,include=T,fig.width=20,fig.height=32}
cowplot::plot_grid( plotlist =c(Other_list,Other_heatmap_list),ncol=3,rel_heights = c(0.8,1)   ) 
```

```{r}
knitr::knit_exit()
```

#########################################################################################################################################################################
# gsea analysis for hallmark and C2 seperately

```{r,include=F}
set.seed(1234) # keep set seed in the front of fgsea loop otherwise it will change the result
fgsea_list=list()
for (i in GMT_files){
 name=gsub(".*/","",i)
 name2=gsub("symbols.gmt.txt","",name)
 pathway= fgsea::gmtPathways(i)
 
 fgseaRes <- fgsea::fgseaMultilevel(pathway, ranks,
                           minSize=5,
                           maxSize=500
                           )
 fgsea_list[[name2]]=fgseaRes

}
Fgsea_all=do.call("rbind",fgsea_list)
Fgsea_all_sig=Fgsea_all%>%dplyr::filter(padj<0.05)%>%arrange(padj)%>%dplyr::rename(gene_counts=size)

```


## Hallmark gene sets
```{r,include=T,fig.width=18,fig.height=8}
Hallmark_pathway= Fgsea_all_sig%>%dplyr::filter(grepl("HALLMARK",pathway))%>%.[,c(1,3,6,7)]%>%arrange(desc(NES))
knitr::kable(Hallmark_pathway,caption = "sig pathways in Hallmark")
```

## Reactome pathway
```{r,include=T}

C2_pathway= Fgsea_all_sig%>%dplyr::filter(!grepl("HALLMARK",pathway))%>%.[,c(1,3,6,7)]%>%arrange(desc(NES))
```


```{r}
dir(Process_data)
write.csv(Hallmark_pathway[,c(1,2,3,4)],file.path(Process_data,"sig_Hallmark_pathway_pathways_in_MPNST_vs_PN.csv"),row.names = F)
write.csv(C2_pathway[,c(1,2,3,4)],file.path(Process_data,"sig_C2_pathway_pathways_in_MPNST_vs_PN.csv"),row.names = F)

```


```{r,include=F,fig.width=18,fig.height=8}
Hallmark_pathway$pathway=gsub("HALLMARK_","",Hallmark_pathway$pathway)
Hallmark_pathway$pathway=gsub("_"," ",Hallmark_pathway$pathway)
Hallmark_pathway$pathway=stringr::str_to_title(Hallmark_pathway$pathway)
Hallmark_pathway$pathway=gsub("G2m","G2M",Hallmark_pathway$pathway)

#P3=Enrich_dotplot(Hallmark_pathway[1:6,],top_number=11,size_max = 10)

Fig_title="EPITHELIAL MESENCHYMAL TRANSITION"%>%stringr::str_to_title()
P4=Volcano_plot_rnaseq_add_label_update_2(MPNST_vs_PN_res_2,pathway_all,"HALLMARK_EPITHELIAL_MESENCHYMAL_TRANSITION",top_gene_number = 10,New_title = Fig_title )

C2_pathway$pathway=gsub("_"," ",C2_pathway$pathway)
#C2_pathway$pathway=stringr::str_to_title(C2_pathway$pathway)
#C2_pathway$pathway=gsub("G2m","G2M",C2_pathway$pathway)
#P5=Enrich_dotplot(C2_pathway[1:10,],11,size_max=10)

# for answering the reviewer's question
Pathway_name=c("MEISSNER_BRAIN_HCP_WITH_H3K27ME3")
Fig_title=gsub("_", " ",Pathway_name)
Fgsea_all%>%dplyr::filter(pathway%in%c("HALLMARK_EPITHELIAL_MESENCHYMAL_TRANSITION", "MEISSNER_BRAIN_HCP_WITH_H3K27ME3"))

Gene_list=pathway_all[Pathway_name]%>%unlist()%>%as.character()

fgsea::plotEnrichment(Gene_list, stats = ranks)
P6=Volcano_plot_rnaseq_add_label_update_2(MPNST_vs_PN_res_2,pathway_all,Pathway_name,top_gene_number = 10,New_title = Fig_title )

# NOT INCLUDE IN THE FINAL VERSION
Other_list=list()
# Reactome_pathway_2=Reactome_pathway%>%dplyr::filter(pathway%in%c("REACTOME_DEGRADATION_OF_THE_EXTRACELLULAR_MATRIX","REACTOME_EXTRACELLULAR_MATRIX_ORGANIZATION","REACTOME_CELL_CYCLE_MITOTIC","REACTOME_EGR2_AND_SOX10_MEDIATED_INITIATION_OF_SCHWANN_CELL_MYELINATION"))
Pathway_name=c("MEISSNER_BRAIN_HCP_WITH_H3K27ME3")
Gene_list=pathway_all[Pathway_name]
 
#P5=Singscore_boxplot(Gene_list,Expression_vst_MPNST_PN_q005,RNAseq_anno,"Tumor_type",Stat_method,Tumor_color2[1:2])
Gene_list2=as.character(unlist(Gene_list))
Gene_list3=intersect( rownames(MPNST_vs_PN_res_q005),Gene_list2)
```

\newpage 

```{r,include=F,fig.width=32,fig.height=18,out.width="40%"}
#gridExtra:: grid.arrange(P1,P2, P3[[1]][[4]], nrow = 2)
#cowplot:: plot_grid(P3,NULL,P4[[1]][[4]],NULL,  nrow = 2,labels = c("C","","D"),rel_widths = c(0.5,1,1,0.1))
#cowplot:: plot_grid(P3,NULL,labels = c("C","V"),nrow=1,scale =c(1,1),label_size = 20,vjust=1)

#gridExtra:: grid.arrange(P1,P2, P3[[1]][[4]], nrow = 2)

#cowplot:: plot_grid(P3,NULL,P4[[1]][[4]],NULL,  nrow = 2,labels = c("C","","D"),rel_widths = c(0.5,1,1,0.1))



#cowplot:: plot_grid(P3,P5,labels = c("C","E"),nrow=1,rel_widths =c(1,1),label_size = 50,vjust=1)
```

```{r}
C2_select=C2_pathway[1:10,]%>%mutate(Geneset="C2_Curated gene sets")
Hallmark_select=Hallmark_pathway[1:6,]%>%mutate(Geneset="Hallmark gene sets")
Merge=rbind(Hallmark_select,C2_select)
Merge$Geneset=factor(Merge$Geneset,c("Hallmark gene sets","C2_Curated gene sets"))
P3=Enrich_dotplot(Merge,30,size_max=15)+facet_wrap(~Geneset,scales = "free")+theme(
    strip.background = element_blank(),                # remove gray background
    strip.text = element_text(size = 50, face = "bold")  # control title text size
  )
```


```{r,include=T,fig.width=40,fig.height=20,out.width="90%"}
# C AND D
P3
```

```{r,include=F,fig.width=32,fig.height=18,out.width="40%"}
#gridExtra:: grid.arrange(P1,P2, P3[[1]][[4]], nrow = 2)

#cowplot:: plot_grid(P3,NULL,P4[[1]][[4]],NULL,  nrow = 2,labels = c("C","","D"),rel_widths = c(0.5,1,1,0.1))
#cowplot:: plot_grid(P4,grid::nullGrob(),labels = c("D",""),nrow=1,rel_widths =c(1,1),label_size = 50,vjust=1)

```

```{r,include=T,fig.width=32,fig.height=18,out.width="70%"}
#gridExtra:: grid.arrange(P1,P2, P3[[1]][[4]], nrow = 2)

#cowplot:: plot_grid(P3,NULL,P4[[1]][[4]],NULL,  nrow = 2,labels = c("C","","D"),rel_widths = c(0.5,1,1,0.1))
cowplot:: plot_grid(P4,P6,labels = c("E","F"),nrow=1,rel_widths =c(1,1),label_size = 50,vjust=1)

```




```{r,include=T,fig.width=32,fig.height=8,out.width="50%"}
#gridExtra:: grid.arrange(P1,P2, P3[[1]][[4]], nrow = 2)

#cowplot:: plot_grid(P3,NULL,P4[[1]][[4]],NULL,  nrow = 2,labels = c("C","","D"),rel_widths = c(0.5,1,1,0.1))
#cowplot:: plot_grid(P6,grid::nullGrob(),labels = c("F",""),nrow=1,rel_widths =c(0.3,1),label_size = 50,vjust=-1)

```

############################################################################################################################################################################################
